---
title: "Spectral Instrument Comparison -- Toolik 2019"
author: "Ruby An"
date: "July 1, 2019"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

R markdown document for comparing spectral instruments at Toolik. Will update with methods eventually.... currently function mainly to organize code into digestable chunks. See plots at the bottom for a "comparison" starting point.

## Required Packages & Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


## Required Packages
library("tidyverse")
library("knitr")

data_path <- "C:/Users/toolik/Desktop/UnispecR/UnispecProtocol/spectral_instrument_comparison/"
data_path <- "/home/ruby/UnispecR/UnispecProtocol/spectral_instrument_comparison/"


band_defns <- tribble(
  ~definition, ~color, ~min, ~max,
  "ITEX", "red", 560, 600,
  "ITEX", "nir", 725, 1000,
  "MODIS", "red", 620, 670, 
  "MODIS", "nir", 841, 876,
  "MODIS", "blue", 459,479,
  "SKYE", "red", 620, 680,
  "SKYE", "nir", 830, 880,
  "SKYE", "blue", 455, 480,
  "Greenseeker", "red", 656, 656,
  "Greenseeker", "nir", 774, 774, 
  "RapidSCAN", "red", 670, 670,
  "RapidSCAN", "nir", 780, 780,
  "RapidSCAN", "red_edge", 730, 730, 
  "ToolikGIS_Drone_2018", "red", 640, 680,
  "ToolikGIS_Drone_2018", "nir", 820, 890,
  "ToolikGIS_MicaSense_2019", "blue", 455, 495,
  "ToolikGIS_MicaSense_2019", "green", 540, 580,
  "ToolikGIS_MicaSense_2019", "red", 658, 678,
  "ToolikGIS_MicaSense_2019", "red_edge", 707, 727,
  "ToolikGIS_MicaSense_2019", "nir", 800, 880,
  "ToolikEDC", "red", 560, 680,
  "ToolikEDC", "nir", 725, 1000
)




```

### Functions 
R functions scraped from Ruby's unispec code to facilitate processing.
```{r, echo = F}
read_spu_file_metadata <- function(filename) {
  
  ruby_year <- str_detect(filename, "2017|2018|2019")
  
  if (ruby_year) {
    ## Reads a .spu file from 2017 or 2018 
    
    # Filename metadata
    filename_metadata <- unlist(str_split(filename, pattern = "/")) %>% last() %>% str_split("_") %>% unlist()
    Site <- filename_metadata[2]
    Date <- filename_metadata[1]
    
    FileNum <- as.integer(str_extract(filename_metadata[3], "\\d{5}")) # extract 5 digits
    
    # Extract info from the file itself, reading metadata from first 9 lines. Create a dataframe
    text <- read_lines(filename, n_max=9)
    DateTime <-  lubridate::mdy_hms(text[3], tz="America/Anchorage")
    Date <- lubridate::date(DateTime)
    Integration_ms <- as.numeric(strsplit(text[8], split = " ")[[1]][3])
    Temp <- as.numeric(strsplit(strsplit(text[5], split = " ")[[1]][4], split="=")[[1]][2])
    Remarks <- text[2]
    
    # Truncated Filename 
    Spufilename <- unlist(str_split(filename, pattern = "/")) %>% last()
    
    # Metadata 
    metaData <- tibble(Site=Site, FileNum=FileNum, Date=Date, DateTime=DateTime, Integration_ms=Integration_ms,
                       Temp=Temp, Remarks=Remarks, spu_filename=Spufilename )
    
  } else {
    ## Reads a generic .spu file. Written for Historic (pre-2017) years. 
    
    # Extract metadata from filenames that have format "DATE/SITE_FILENUM.spu", e.g. "2018-06-22/DHT_00000.spu"
    Site <- toupper(str_replace(filename,".*[/](.*)[_]+.*$","\\1")) # get string after / & before _
    
    # Extract metadata from filenames that have format "DATESITEFILENUM.spu", e.g. "JUN8LOF100036.spu"
    if (str_length(Site) > 5) {
      Site <- toupper(str_replace(filename,"(^.*?\\d{1,2})\\s*([a-zA-Z]*)(\\d{5,7}\\.spu$)","\\2"))
      # For 2012 and 2013 the spu filenames have ddmmmsite format; need to remove the 3 letter month.
      if (str_length(Site)> 5){
        pattern <- c("MAY","JUN","JUL", "AUG")
        for (i in 1:4){Site<- sub(pattern[i], "", Site)}
      }
    }
    
    # Avoid issues of digits in site/block name reading in as part of FileNum
    ## Based on the Unispec DC using 5 digits for automatic file numbering
    FileNum <- as.integer(str_replace(filename,"(^.*?\\d{1,2})(\\D*)(\\d{5})(\\.spu$)","\\3"))
    
    # Extract info from the file itself, reading metadata from first 9 lines. Create a dataframe
    text <- read_lines(filename, n_max=9)
    DateTime <-  lubridate::mdy_hms(text[3], tz="America/Anchorage")
    Date <- lubridate::date(DateTime)
    Integration_ms <- as.numeric(strsplit(text[8], split = " ")[[1]][3])
    Temp <- as.numeric(strsplit(strsplit(text[5], split = " ")[[1]][4], split="=")[[1]][2])
    Remarks <- text[2]
    
    #Extract the file name in the spu file as a check. Some file names have spaces 
    Spufilename <- tolower(str_replace(text[1],".*[\\\\]([A-z0-9.]\\s*)","\\1")) 
    Spufilename <- str_replace(Spufilename,"\"","") # remove trailing quote
    #Extract 
    Spufilename_file <- tolower(str_replace(filename,".*[\\\\/]([A-z0-9.]\\s*)","\\1"))
    
    
    
    metaData <- tibble(Site=Site, FileNum=FileNum, Date=Date, DateTime=DateTime, Integration_ms=Integration_ms,
                       Temp=Temp, Remarks=Remarks, spu_filename=Spufilename )
  }

  print(Spufilename) # use for error checking
  
  return(metaData)
}

read_spu_file_spectra <- function(filename) {
  # For a generic .spu file regardless of name
  
  # Read spectral intensity data into dataframe
  data <- read.table(file = filename, skip = 9, col.names = c("Wavelength", "ChB", "ChA")) %>% 
    mutate(Reflectance = ChB/ChA) 
  
  # Only use the valid spectra
  #  data_valid <-  subset(data, Wavelength > 400 & Wavelength < 1000) 
  
  return(data)
}

interpolate_spectra <- function(spectra) {
  # interpolates unispec spectra to 1nm 
  # input: dataframe with Wavelength & Reflectance columns
  # output: dataframe with Wavelength (every 1 nm) & Reflectance columns 
  #spectra <- spectra[[1]]
  
  is.null(spectra)
  interp <- approx(x = spectra$Wavelength, y = spectra$Reflectance, 
                   xout = seq(400,1000, by = 1), method = "linear")
  spectra_interp <- tibble(Wavelength = interp$x, Reflectance = interp$y)
  
  return(spectra_interp)
}

calculate_spectral_bands <- function(spectra, band_defns, instruments) {
  # Calculates spectral bands from dataframe including Wavelength & Reflectance
  ## inputs: spectra - Wavelength, Reflectance columns
  ##         band_defns : wavelengths definining colors 
  ##         instrument : e.g. MODIS, SKYE, ITEX
  ##         bands   : the spectral bands to return, e.g. red, blue, nir, etc. 
  ## output: spectra_bands = dataframe with Definition, Band, Averaged Reflectance
  

  bands <- band_defns %>% 
    filter(definition %in% instruments)
  
  # vector of wavelengths, one set per instrument
  wavelengths <- seq(300, 1500, by = 1)
  
  # dataframe of wavelengths labeled by instrument & color
  bands_df <- tibble(Wavelength = rep(wavelengths, times = length(instruments)), 
                     definition = rep(instruments, each = length(wavelengths))) %>% 
    full_join(bands) %>% 
    mutate(color_match = ifelse(Wavelength >= min & Wavelength <= max, color, NA)) %>% 
    select(Wavelength, definition, color_match) %>%
    distinct()
  
  ## DATA: join to measured spectra 
  spectra_bands <- full_join(spectra, bands_df) %>% 
    group_by(definition, color_match) %>% 
    summarize(average_reflectance = mean(Reflectance)) %>% 
    filter(!is.na(color_match)) %>% 
    rename(band = color_match)
  
  return(spectra_bands) 
}

plot_spectra <- function(df_subset) {
  # Plot spectra from a subset of a dataframe
  
  plot_check <- df_subset %>%
    unnest(Spectra) %>%
    filter(Wavelength > 400, Wavelength < 1000) %>% 
    gather(key = Channel, value = Intensity, ChB, ChA) %>%
    gather(key = ref_part, value = Reflectance_Intensity, Intensity, Reflectance)
  
  
  ## Plot Specified Correction Factors for quality check
  plot_zoom <- ggplot(data = plot_check, mapping = aes(x = Wavelength, y = Reflectance_Intensity)) +
    geom_line(aes(color=Channel)) +
    facet_grid(ref_part ~ spu_filename, scales = "free")
  # 
  # if("Treatment" %in% names(df_subset)) { # use for datafarmes with and without metadata
  #   plot_zoom <- plot_zoom + 
  #     facet_grid(ref_part ~ DateTime + Site + FileNum + Treatment, scales="free")
  #   
  # } else {
  #   plot_zoom <- plot_zoom + 
  #     facet_grid(ref_part ~ DateTime + Site + FileNum, scales="free")
  #   
  # }
  
  return(plot_zoom)
}



```

## Load Data & File Keys

###  Unispec-DC Pre-Processing 
Function for correcting a set of files with a set of listed references. This is the "multispec" equivalent for Unispec-DC files.
```{r, echo = FALSE, message = F, eval = F}

### CODE DEVELOPMENT FOR CORRECTING UNISPEC-DC FILE SUBSETS MANUALLY 

## File Key (label scans) 
unispec_dc_filekey <- read_csv(paste0(data_path, "2019-06-26_NDVI-comparison_Unispec-DC_filekey.csv")) 

spu_filekey <- unispec_dc_filekey 




## Read-in all files 
spu_files <- list.files(path = data_path, pattern = ".spu$", full.names = T, recursive=T)

### read spu metadata : info from .spu header (Timestamp, Integration, Temp, Remarks)
spu_metadata <- map_dfr(spu_files, read_spu_file_metadata) %>% 
  mutate(spu_filename_full = spu_files)

### read actual spectra 
spu_data <- spu_metadata %>%
  mutate(Spectra=map(spu_filename_full, function(x) read_spu_file_spectra(x)))




## Standardize Dataframe
unispec_dc_data <- spu_data %>% select(spu_filename, Site, FileNum, Date, DateTime, Integration_ms, Temp, Remarks, Spectra, everything()) %>%  ## METADATA
  rename(FILENAME = spu_filename)

unispec_dc_dataframe <- unispec_dc_filekey %>% full_join(unispec_dc_data) %>% 
  mutate(NOTES = as.character(NOTES)) 



## Datascans (not REFs)
### 2 column dataframe: (1) spu_filename; (2) Spectra <list column: dataframe w/Wavelength, ChA, ChB, Reflectance]
raw_spu_data <- unispec_dc_dataframe %>% select(FILENAME, Spectra) %>% 
  rename(spu_filename = FILENAME) %>% slice(7:113) 

## REFscans 
### 2 column dataframe: (1) spu_filename; (2) Spectra <list column: dataframe w/Wavelength, ChA, ChB, Reflectance]
ref_dataframe <- unispec_dc_dataframe %>% filter(str_detect(LOCATION,"REF")) %>% select(FILENAME, Spectra) %>% 
  rename(spu_filename = FILENAME)

## Calculate Correction Factors from REFs
ref_correction <- ref_dataframe %>% unnest(Spectra) %>% 
  mutate(CorrectionFactor = 1/Reflectance) %>% 
  group_by(Wavelength) %>% 
  summarize(CorrectionFactor = mean(CorrectionFactor), ref_filenames = str_c(spu_filename, collapse = ", "))

## Plot REFERENCES to check
ref_dataframe %>% plot_spectra() + ggtitle("REFscan spectra")

## Apply REF Correction Factor to get corrected spectra:

## Corrected Datascans 
### 3 column dataframe:
#     (1) spu_filename
#     (2) ref_filenames (comma-separated string)
#     (3) Spectra <list-column> :
#           5-column tibble w/Wavelength, ChB, ChA, Reflectance, CorrectionFactor
#           "Reflectance" = ChB/ChA*CorrectionFactor

corrected_spu_data <- raw_spu_data %>% unnest(Spectra) %>% 
  left_join(ref_correction) %>% 
  mutate(Reflectance = Reflectance*CorrectionFactor) %>% ## update Reflectance w/ REF correction factor
  nest(Wavelength, ChB, ChA, Reflectance, CorrectionFactor, .key = Spectra)


## Interpolated Datascans : to 1 nm resolution (Multispec Equivalent)
### only eeds "Spectra" column to include "Reflectance" column -- then it will work! >> UPDATE LATER TO INTERPOLATE ANY COLUMN?
unispec_dc_spectra <- corrected_spu_data %>% select(spu_filename, Spectra) %>% 
  rename(FILENAME = spu_filename) %>% ## to join with filekey
  mutate(interpolated_spectra = map(Spectra, .f = interpolate_spectra)) # Spectra must have "Reflectance" column

## SAVE CORRECTED & INTERPOLATED SPECTRA 

### R file format 
write_rds(unispec_dc_spectra, paste0(data_path, "2019-06-26_NDVI-comparison_Unispec-DC_spectra.rds"))

### .csv file format (corrected, interpolated reflectance only)
spu_to_csv_data <-  unispec_dc_spectra %>% unnest(interpolated_spectra)
write_csv(spu_to_csv_data, paste0(data_path, "2019-06-26_NDVI-comparison_Unispec-DC_corrected-interpolated-reflectance.csv"))
```



### Read & Calculate NDVI 

Load all data files.
```{r, echo = F, message =F}

#### RapidSCAN & Greenseeker DATA & KEY -------------
scanners <- read_csv(paste0(data_path, "2019-06-26_NDVI-comparison_RapidSCAN-Greenseeker_data.csv"))

#### Unispec-SC DATA & KEY -------------

## File Key
unispec_sc_filekey <- read_csv(paste0(data_path, "2019-06-26_NDVI-comparison_Unispec-SC_filekey.csv"))

## Data -- corrected with Multispec by EDC
unispec_sc_data <- read_csv(paste0(data_path, "2019-06-26_NDVI-comparison_Unispec-SC_data.csv"), skip = 1, col_names = T) %>% 
  select(-X9) %>% # remove empty column
  rename(Height = `Instrument Height (m)`, FILENAME = `File Name`) %>% 
  gather(key = Wavelength, value = Reflectance, `310`:`1130`) %>%  ## METADATA
  mutate(Wavelength = as.integer(Wavelength))

unispec_sc_spectra <- unispec_sc_data %>% group_by(FILENAME) %>% 
  nest(Wavelength, Reflectance, .key = Spectra) ## to join with filekey

## Calculate NDVI
unispec_sc_dataframe <- unispec_sc_filekey %>% full_join(unispec_sc_spectra) %>% 
  mutate(Bands = map(Spectra, function(x) calculate_spectral_bands(x, band_defns = band_defns, instruments = c("ToolikEDC", "MODIS", "Greenseeker", "RapidSCAN"))))


#### Unispec-DC -------------

## File Key
unispec_dc_filekey <- read_csv(paste0(data_path, "2019-06-26_NDVI-comparison_Unispec-DC_filekey.csv")) 

## Data (.csv saved post R-processing to correct w/White REF & interpolate to 1nm)
unispec_dc_all_data <- read_rds(paste0(data_path, "2019-06-26_NDVI-comparison_Unispec-DC_spectra.rds"))

unispec_dc_spectra <- read_csv(paste0(data_path, "2019-06-26_NDVI-comparison_Unispec-DC_corrected-interpolated-reflectance.csv")) %>% 
  group_by(FILENAME) %>% 
  nest(Wavelength, Reflectance, .key = Spectra) ## to join with filekey

# Calculate Spectral Bands
unispec_dc_dataframe <- unispec_dc_spectra %>% 
  left_join(unispec_dc_filekey)

  
#### ASD -------------

## File Key
asd_filekey <- read_csv(paste0(data_path, "2019-06-26_NDVI-comparison_ASD_filekey.csv"))

## Data 
asd_spectra <- read_csv(paste0(data_path, "2019-06-26_NDVI-comparison_ASD_data.csv"), col_names = T) %>% 
  gather(2:94, key = FILENAME, value = Reflectance) %>% 
  rename(Wavelength = 1) %>% 
  group_by(FILENAME) %>% nest(Wavelength, Reflectance, .key = Spectra)

asd_dataframe <- asd_filekey %>% 
  full_join(asd_spectra) %>%  mutate(NOTES = as.character(NOTES)) %>% 
  mutate(Bands = map(Spectra, function(x) calculate_spectral_bands(x, band_defns = band_defns, instruments = c("ToolikEDC", "MODIS", "Greenseeker", "RapidSCAN"))))

## Calculate NDVI 



```



## Multispectral Comparisons (ASD, Unispec-SC, Unispec-DC)

Instrument comparison on LTER 2006 MAT plots (Block 2, CT [tussock] & F10 [shrub]) and nearby Dry Heath [heath].

```{r pressure, echo=FALSE, message = F}

## Calculate Spectral Bands for Multispectral Instruments
multispectral_df <- bind_rows(unispec_sc_dataframe, unispec_dc_dataframe) %>% 
  bind_rows(asd_dataframe) %>% 
  mutate(Bands = map(Spectra, function(x) calculate_spectral_bands(x, band_defns = band_defns, instruments = c("ToolikEDC", "MODIS", "Greenseeker", "RapidSCAN"))))
  


## Recode LOCATION variable to be Human-Readable 
multispec_df <- multispectral_df %>% mutate(LOCATION_LONG = LOCATION)

Location_Names <- list(`LMAT_B2_CT_EDC_2-1` = "T1",
                       `LMAT_B2_CT_EDC_2-2` = "T2",
                       `LMAT_B2_CT_EDC_2-3` = "T3", 
                       `LMAT_B2_F10_15m` = "S1", 
                       `LMAT_B2_F10_10m` = "S2",
                       `LMAT_B2_F10_5m` = "S3", 
                       `ITEX_BOARDWALK-1` = "H1",
                       `ITEX_BOARDWALK-2` = "H2",
                       `ITEX_BOARDWALK-3` = "H3")

multispec_df$LOCATION <- recode (multispec_df$LOCATION_LONG, !!!Location_Names, .default = multispec_df$LOCATION)

scanners$LOCATION <- recode(scanners$LOCATION, !!!Location_Names, .default = scanners$LOCATION)

# Write this to .rds dataframe to read-in later 
multispec_bands_name <- "UnispecProtocol/spectral_instrument_comparison/spec_synthesis_multispec_band_df.rds"
write_rds(multispec_df, multispec_bands_name)

```

### Plot full spectra

Plots full spectra for ASD and Unispec's. 
```{r, echo = F, message = F}

multispec_df <- read_rds(multispec_bands_name)

multispec_df_plot <- multispec_df %>% unnest(Spectra) %>% 
  filter(nchar(LOCATION) == 2) %>% # exclude whole plot Unispec-DC and ASD measurements
  filter(Wavelength > 300, Wavelength < 1000)

ggplot(data = multispec_df_plot, aes(x = Wavelength, y = Reflectance)) +
  geom_line(aes(color = INSTRUMENT, group = FILENAME)) +
  facet_wrap(vars(LOCATION))

```

### NDVI comparisons
Columns are "NDVI-definition" and rows are vegetation type. 
```{r, echo=F, message=F}

scanners_df <- scanners %>% 
  mutate(definition = ifelse(INSTRUMENT == "RapidSCAN", "RapidSCAN", "Greenseeker"))

ndvi_df <- multispec_df %>% unnest(Bands) %>% 
  #select(INSTRUMENT, VEGETATION, LOCATION, REPLICATE, FILENAME, definition, band, average_reflectance) %>% 
  filter(nchar(LOCATION) == 2) %>% # exclude whole plot Unispec-DC and ASD measurements
  ## Calculate NDVI
  spread(band, average_reflectance) %>% 
  mutate(NDVI = (nir - red) / (nir + red)) %>% 
  ## Add Greenseeker & RapidSCAN NDVI data
  bind_rows(scanners_df) %>% 
  # Exclude one recording mistake point (outlier)
  filter(!(VEGETATION == "Tussock" & REPLICATE == 4 & INSTRUMENT == "RapidSCAN")) %>% 
  ## Remove 1.5m Unispec-SC measurements (>1m2 footprint, based on 80cm diameter footprint at 1m height)
  rename(SC_HEIGHT = `Instrument Height (m)`) %>% 
  filter(SC_HEIGHT == 1 | is.na(SC_HEIGHT)) 
  
## PLOT IT
ggplot(ndvi_df, aes(x=INSTRUMENT, y = NDVI)) +
  geom_boxplot(aes(fill=INSTRUMENT)) +
  facet_grid(VEGETATION ~ definition) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

## Save NDVI data
# ndvi_to_save <- ndvi_df %>% 
#   select(INSTRUMENT, VEGETATION, LOCATION, REPLICATE, SC_HEIGHT, FILENAME, NOTES, definition, NDVI)
# 
# write_csv(ndvi_to_save, paste0(data_path, "2019-06-26_NDVI-compare_ndvi-data.csv"))

```

### METHOD comp: Unispec-DC vs ASD
```{r}

```


## Seasonal Trend Comparison (2017, 2018, 2019)

Comparison between EDC Unispec-SC and LTER Unispec-DC data. 

```{r}
## Load processed data, currently for Toolik RA computer only 
hst_data_path <- "/Users/toolik/OneDrive - Marine Biological Laboratory/Toolik Terrestrial/UnispecData/"


## 2017 
spu_2017 <- read_rds(paste0(hst_data_path, "2017_processed_spu_data.rds"))
key_2017 <- read_csv(paste0(hst_data_path, "2017_unispec_file_key.csv"))


df_2017 <- spu_2017 %>% full_join(key_2017) %>%
  ## LMAT & HTH Sites 
  filter(str_detect(Site, "LMAT|HTH")) %>% 
  ## CT Treatments only 
  filter(Treatment %in% c("CT", "CT1", "CT2")) %>% 
  ## FORMAT CORRECTLY
  mutate(Replicate = as.character(Replicate))


## 2018
spu_2018 <- read_rds(paste0(hst_data_path, "2018_unispec_dataframe_clean.rds")) %>% 
  ## LMAT & HTH Sites 
  filter(str_detect(Site, "LMAT|HTH")) %>% 
  ## CT Treatments only 
  filter(Treatment %in% c("CT", "CT1", "CT2"))

raw_2018 <- df_2018 %>% unnest(raw_spectra) %>% rename(raw_reflectance = Reflectance)
corrected_2018 <- df_2018 %>% unnest(corrected_spectra) %>% 
  rename(corrected_reflectance = Reflectance) %>% select(-ChA, -ChB) 

df_2018 <- full_join(raw_2018, corrected_2018) %>% 
  mutate(CorrectionFactor = corrected_reflectance/raw_reflectance) %>% 
  nest(.key = Spectra, Wavelength, ChA, ChB, raw_reflectance, CorrectionFactor, corrected_reflectance)

## 2019
spu_2019 <- read_rds(paste0(data_path, "2019_processed_spu_data.rds"))
key_2019 <- read_csv(paste0(data_path, "2019_unispec_file_key.csv"))

df_2019 <- full_join(spu_2019, key_2019) %>% 
  ## LMAT & HTH Sites 
  filter(str_detect(Site, "LMAT|DHT")) %>% 
  ## CT Treatments only 
  filter(Treatment %in% c("CT", "CT1", "CT2")) %>% 
  ## FORMAT CORRECTLY
  mutate(Replicate = as.character(Replicate))
  


## COMBINED
df <- bind_rows(df_2017, df_2018) %>% bind_rows(df_2019) %>% 
  unnest(Spectra)

write_csv(df, paste0(data_path, "2017-2019_LMAT-DHT_CT_unispec-dc_spu_data.csv"))

```














































