---
title: 'Rapidscan vs Unispec-DC: 2019-06-02'
author: "Ruby An"
date: "December 19, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require("knitr")
source("UnispecProtocol/unispec_protocol_functions.R") # file loads required functions

interpolate_spectra <- function(spectra) {
  # interpolates unispec spectra to 1nm 
  # input: dataframe with Wavelength & Reflectance columns
  # output: dataframe with Wavelength (every 1 nm) & Reflectance columns 
  #spectra <- spectra[[1]]
  
  is.null(spectra)
  interp <- approx(x = spectra$Wavelength, y = spectra$Reflectance, 
                   xout = seq(400,1000, by = 1), method = "linear")
  spectra_interp <- tibble(Wavelength = interp$x, Reflectance = interp$y)
  
  return(spectra_interp)
}
```


## Overview 
These measurements were taken to (1) compare Unispec-DC, RapidSCAN, and Greenseeker NDVI measurements and (2) assess plot heterogeneity early season. 

## Description of Sampling 


- UNISPEC METHOD: unispec measurements taken in normal direction (downslope to upslope) at normal 5 points of LTER measurements + 4 extra measurements between (25 measurements total, except when mistakes made, see Notes column). Thes 5 points are marked by silver tags on the boardwalk, and are typically located at the 10-15m location of the plot on the North side towards the lake (measured downslope to upslope). Measurements are taken 0.5 m into the plot (length of unispec monopod arm).

- GREENSEEKER METHOD: one circular motion rotation at same height/area as unispec. Swipe is taken by holding down the trigger and walking along the 5m transect; 

- RAPIDSCAN METHOD: held in same location as unispec downfacing foreoptic, scan for approx. 3 seconds. Swipe taken by walking with RapidSCAN along 5m transect (silver tags). 

```{r, echo = F}
band_defns <- tribble(
  ~definition, ~color, ~min, ~max,
  "ITEX", "red", 560, 600,
  "ITEX", "nir", 725, 1000,
  "MODIS", "red", 620, 670, 
  "MODIS", "nir", 841, 876,
  "MODIS", "blue", 459,479,
  "SKYE", "red", 620, 680,
  "SKYE", "nir", 830, 880,
  "SKYE", "blue", 455, 480,
  "Greenseeker", "red", 656, 656,
  "Greenseeker", "nir", 774, 774, 
  "RapidSCAN", "red", 670, 670,
  "RapidSCAN", "nir", 780, 780,
  "RapidSCAN", "red_edge", 730, 730, 
  "ToolikGIS_Drone_2018", "red", 640, 680,
  "ToolikGIS_Drone_2018", "nir", 820, 890,
  "ToolikGIS_MicaSense_2019", "blue", 455, 495,
  "ToolikGIS_MicaSense_2019", "green", 540, 580,
  "ToolikGIS_MicaSense_2019", "red", 658, 678,
  "ToolikGIS_MicaSense_2019", "red_edge", 707, 727,
  "ToolikGIS_MicaSense_2019", "nir", 800, 880,
  "ToolikEDC", "red", 560, 680,
  "ToolikEDC", "nir", 725, 1000
)

kable(band_defns)
```


## Data Summary  
This section includes hidden code chunks to undertake preliminary processing of data. If of interest, see R Markdown document. 

The two relevant dataframes this produces are:

1. "2019-06-02_scanner_ndvi.csv". This includes Greenseeker and Rapidscan data. 
2. "2019-06-02_unispec_ndvi.csv". This includes Unispec-DC NDVI calculated using MODIS, Rapidscan, and Greenseeker band definitions (see above). 

```{r, read_in_data, include=F }

# Set directory 
path <- "spectral_instrument_comparison/2019-06-02/"

# >>>>>>>>>>>>>>>>>> Greenseeker and RapidSCAN data 
scanner_data <- read_csv("spectral_instrument_comparison/2019-06-02/2019-06-02_LMAT_spectral-instrument-comparison.csv")

write_csv(scanner_data, "spectral_instrument_comparison/2019-06-02/2019-06-02_scanner_ndvi.csv")


# >>>>>>>>>>>>>>>>>> UNISPEC DATA 
unispec_file_key <- read_csv("spectral_instrument_comparison/2019-06-02/2019-06-02_LMAT_unispec_file_key.csv")

# Read .spu raw files
raw_data_path <- paste0(path, "raw_spu_files")
spu_files <- list.files(path = raw_data_path, pattern = ".spu$", full.names = T, recursive=T)

# Read spu_metadata (integration time, temp, datetime, etc.) 
spu_metadata <- map_dfr(spu_files, read_spu_file_metadata) %>% 
  mutate(spu_filename_full = spu_files)

# Read spectral data ==> Dataframe of raw spp spectra
spu_data <- spu_metadata %>% 
  # Read in spectral data
  mutate(Spectra=map(spu_filename_full, function(x) read_spu_file_spectra(x))) %>% 
  # Order columns 
  select(spu_filename, Site, FileNum, Date, DateTime, Integration_ms, Temp, Remarks, Spectra, everything())

## Join file metadata to spu data 
raw_data <- full_join(unispec_file_key, spu_data)

# Select references 
ref_data <- raw_data %>% filter(Treatment == "REF")

## Average 5 chosen ref measurements per DATE/SITE/BLOCK 
ref_summary <- ref_data %>% 
  ## The following steps expand the "Block" column to create one REF set per Block per Site. This structure is necessary for situtations where different refs are used for different blocks at the same site. 
  separate(Block, into = c("BX1", "BX2", "BX3", "BX4"), sep = ",") %>% #1: expand string entry in "Block" into separate columns -- NOTE: this step throws a "Warning: Expected 4 pieces." for sites w/less than 4 blocks
  gather(Block, BlockString, BX1:BX4) %>% #2: re-condense into one column, generates correct number of rows per site AND per block
  mutate(Block = str_squish(BlockString), BlockString=NULL) %>% #3: replace placeholder column names w/"B1-B4". Also removes whitespace from BlockString contents introduced by "separate" function
  filter(!is.na(Block)) %>% #4: remove empty rows for sites w/out B3 or B4
  ### Unnest Spectra & calculate
  unnest(Spectra) %>% 
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  mutate(CorrectionFactor = 1/Reflectance) %>% 
  ### The following code group repeated REF measurements, and takes the mean 
  group_by(Date,Site,Block,Wavelength) %>% 
  # group_by(Date,Site,Block,Wavelength, Integration_ms) %>% # to separate integration times
  summarize(ChA_REF = mean(ChA), ChB_REF = mean(ChB), CorrectionFactor = mean(ChA/ChB), int_REF = mean(Integration_ms), Notes_REF = str_c(Notes, collapse = "; "), ref_filenames = str_c(spu_filename,collapse = ", "))

# Apply references to raw spu data ==> Dataframe of corrected spu spectra 
corrected_data <- raw_data %>% 
  filter(!str_detect(Treatment, "REF|DARK")) %>% 
  unnest(Spectra) %>% 
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  left_join(ref_summary) %>% 
  select(Date, DateTime, Site, Block, Treatment, Replicate, spu_filename, FileNum, Integration_ms, Weather, Notes, Notes_REF, ref_filenames, int_REF, Wavelength, ChB, ChA, ChB_REF, ChA_REF, CorrectionFactor) %>%
  mutate(raw_reflectance = ChB/ChA) %>% # the raw reflectance
  mutate(Reflectance = raw_reflectance*CorrectionFactor) %>%  # the corrected reflectance
  nest(Wavelength:Reflectance, .key = Spectra) %>% 
  # interpolate Specta to 1nm 
  mutate(interpolated_spectra = map(Spectra, .f = interpolate_spectra)) # Spectra must have "Reflectance" column

# Calculate Spectral Bands 
band_data <- corrected_data %>% 
  # requires interpolated spectra
  mutate(Bands = map(interpolated_spectra, function(x) calculate_spectral_bands(x, band_defns = band_defns, instrument = c("MODIS", "Greenseeker", "RapidSCAN")))) 

# Save spectral band data to load later 
write_rds(band_data, path = "spectral_instrument_comparison/2019-06-02/2019-06-02_unispec_spectral_bands.rds")

# Calculate NDVI data 
ndvi_data <- band_data %>% unnest(Bands) %>% 
  spread(key = band, value = average_reflectance) %>% 
  mutate(NDVI = (nir-red)/(nir+red))


# Save NDVI data
write_csv(ndvi_data %>% select(spu_filename, Date, Site, Block, Treatment, Replicate, definition, NDVI), path = "spectral_instrument_comparison/2019-06-02/2019-06-02_unispec_ndvi.csv")
```



# Quick Plots

This boxplot shows the NDVI per plot. Plot identity (treatment) on the x-axis. NDVI on the y-axis. Please note that the number of samples in each of the bar plots is different, mainly 5 per each Rapidscan and Greenseeker and 20 for the unispec. >> you might consider only including some unispec measurements, etc. real stats have not been done! 
```{r, echo=F}

d1 <- ndvi_data %>% mutate(Replicate = as.character(Replicate)) %>% 
  select(Date, Site, Block, Treatment, Replicate, definition, NDVI, Notes) %>% 
  mutate(instrument = "Unispec-DC") %>% 
  filter(definition == "RapidSCAN") # only rapidscan definition considered, can alter to also look at greenseeker 

d2 <- scanner_data %>% 
  select(-spu_filenames, -Unispec_NDVI) %>% 
  rename(RapidSCAN = Rapidscan) %>% 
  gather(RapidSCAN, Greenseeker, key = definition, value = NDVI) %>% 
  mutate(instrument = definition) %>% 
  filter(definition == "RapidSCAN")

ndvi_df <- rbind(d1, d2)

  
## PLOT IT
ggplot(ndvi_df, aes(x=Treatment, y = NDVI)) +
  geom_boxplot(aes(fill=instrument)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


