---
title: "2019 Unispec Data Investigations"
author: "Ruby An"
date: "June 24, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(lubridate)
require(knitr)

```

## Load Data 

2019 unispec data: updated as summer progresses 

```{r load_data, echo=F, include =F}

## MANUAL
data_path <-  "/Users/toolik/OneDrive - Marine Biological Laboratory/Toolik Terrestrial/UnispecData/2019/"
data_path <- "/home/ruby/UnispecData/2019/"
dir_year <- "2019"

## INTERACTIVE

## Raw Data
raw_filename <- paste0(data_path, dir_year, "_raw_spectra.rds")
spu_spectra <- read_rds(raw_filename)

raw_keyname <- paste0(data_path, dir_year, "_raw_spectra_file_metadata.csv")
spu_key <- read_csv(raw_keyname)

spu_dataframe <- full_join(spu_spectra, spu_key) %>% 
    #mutate(DateTime = lubridate::ymd_hms(DateTime, tz = "US/Alaska")) %>% 
  mutate(Date = str_extract(spu_filename, pattern = "^[0-9]{4}-[0-9]{2}-[0-9]{2}")) %>%  # from filename, not DateTime (sometimes discrepancy in unispec instrument time)
arrange(DateTime)

## Processed Data 
processed_dataframe <- read_rds(paste0(data_path, "2019_processed_spectra.rds")) %>% 
  mutate(DateTime = lubridate::as_datetime(DateTime, tz="US/Alaska")) 

## Index Data
index_data <- read_csv(paste0(data_path, "2019_processed_veg-indices.csv"))

ndvi_data <- index_data %>%
  select(-BandDefinition) %>% 
  spread(Index, Value) %>% select(-EVI, -EVI2)

## File Key
unispec_file_key <- read_csv(paste0(data_path, "2019_processed_unispec_file_key.csv")) 

## joined data 
df <- unispec_file_key %>% left_join(processed_dataframe)
```

```{r, echo=F}
## Summary table: Sites per Date
df %>% group_by(Date) %>% 
  filter(!is.na(Treatment)) %>% 
  filter(!str_detect(Site, "_")) %>% 
  summarize(Sites = str_c(str_replace_na(unique(Site)), collapse = ", "), Num_Files = n()) %>% kable

# Dates per Site
df %>% group_by(Site, Date) %>% 
  summarize(Num_files = n()) %>% print(n=50)

```


## Plot Heterogeneity?  

Comparing normal subset of 5 measurements along 5m to 15 measurements along 20m (or 10 measurements along 10m for smaller plots, e.g. WSG).

### Difference in Mean & Variance

```{r, echo=F, message=F}
check_trtmts <- c("CT", "F0.5", "F1", "F2", "F5", "F10", "N", "P", "NP")  


all_ndvi <- unispec_file_key %>% 
  # filter(as.character(Date) %in% c("2019-06-21", "2019-06-22", "2019-06-18", "2019-06-19", "2019-06-20")) %>%
  filter(!str_detect(Treatment, "DARK|REF|EX")) %>% 
  filter(Treatment %in% check_trtmts) %>% 
  mutate(Site = ifelse(Site %in% c("WSG1", "WSG23"), "WSG", Site))  %>% 
  filter(!str_detect(Block, "EDC")) %>% 
  filter(Block %in% c("B1", "B2", "B3", "B4")) %>% 
  left_join(ndvi_data) %>% 
  mutate(measurement_area = "full_side")

tagged_ndvi <- all_ndvi %>% 
  filter(str_detect(ScanLocation, pattern = "[0-9]S")) %>% 
  mutate(measurement_area = "tagged_transect")

## Summary tables of FileNumbers 
tagged_ndvi %>% 
  group_by(Site, Date, Block, Treatment) %>% 
  summarize(Num_files = n()) %>% 
  print(n=50) 

all_ndvi %>%  group_by(Site, Date) %>%  
  summarize(Num_files = n()) %>% print(n=50)

## Differences in Means / Standard Deviation
all_means <- all_ndvi %>% 
  group_by(Date, Site, Block, Treatment) %>% 
  summarise_at(vars(NDVI), list(mean = ~mean(.), sd = ~sd(.), se = ~sd(.)/sqrt(n()))) 
  
tagged_means <- tagged_ndvi %>% 
  group_by(Date, Site, Block, Treatment) %>% 
  summarise_at(.vars = vars(NDVI), list(mean = ~mean(.), sd = ~sd(.), se= ~sd(.)/sqrt(n()))) %>% 
  rename(mean_tagged = mean, sd_tagged = sd, se_tagged = se)
  
difference_means <- right_join(all_means, tagged_means) %>% 
  mutate(diff_means = mean - mean_tagged, diff_sd = sd - sd_tagged, diff_se = se - se_tagged) %>% 
    mutate(Treatment = ifelse(Treatment %in% c("CT1", "CT2", "CT"), "CT", Treatment)) 



## Plot options
np_colors <- RColorBrewer::brewer.pal(5, "YlGnBu")
## PLOT OPTIONS
# Color Palettes 
n_yellow <- rgb(255, 192, 0, maxColorValue = 255)
p_blue <- rgb(46, 117, 182, maxColorValue = 255)
np_green <- rgb(112, 173, 71, maxColorValue = 255) #lmat_colors[5] 
ct_gray <- rgb(175, 171, 171, maxColorValue = 255)

lmat_colors <- c(rgb(226, 240, 217, maxColorValue = 255), rgb(169, 209, 142, maxColorValue = 255), rgb(112, 173, 71, maxColorValue = 255), rgb(84, 130, 53, maxColorValue = 255), rgb(56, 87, 35, maxColorValue = 255))

## PLOTS - means 
ggplot(data = difference_means) + 
  geom_histogram(aes(x = diff_means, fill = Site), position="stack", binwidth = 0.01)  +
  labs(title = "Difference in plot mean NDVI -- by site", x = "Mean NDVI (full plot) - Mean NDVI (tagged 5m transect)") + 
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = c("SHB" = "orange4", "HIST" = "darkgreen", "MAT" = "green4", "LMAT" = "chartreuse3", "WSG" = "dodgerblue2", "MNAT" = "darkorchid", "NANT" = "mediumpurple4", "DHT" = "firebrick")) + facet_wrap(vars(Site))

ggplot(data = difference_means) + 
  geom_histogram(aes(x = diff_means, fill = Treatment), position="stack", binwidth = 0.01)  +
  geom_vline( xintercept =  0)  +
  labs(title = "Difference in plot mean NDVI -- by Treatment",  x = "Mean NDVI (full plot) - Mean NDVI (tagged 5m transect)") +
  scale_fill_manual(values=c("CT" = ct_gray, "CT1" = ct_gray, "CT2"= ct_gray, "GH-CT" = ct_gray,
                              "N" = n_yellow, "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = p_blue,
                              "S" = "purple1", "L" = "purple2",
                              "NP" = lmat_colors[5],
                              "F0.5" = lmat_colors[1],
                              "F1" = lmat_colors[2],
                              "F2" = lmat_colors[3],
                              "F5" = lmat_colors[4],
                              "F10" = lmat_colors[5])) +
  facet_wrap(vars(Site))


### PLOTS 

ggplot(data = difference_means) + 
  geom_histogram(aes(x = diff_se, fill = Site), position="stack", binwidth = 0.01)  +
  labs(title = "Difference in standard error of NDVI -- by site",  x = "se NDVI (full plot) - se NDVI (tagged 5m transect)") + 
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = c("SHB" = "orange4", "HIST" = "darkgreen", "MAT" = "green4", "LMAT" = "chartreuse3", "WSG" = "dodgerblue2", "MNAT" = "darkorchid", "NANT" = "mediumpurple4", "DHT" = "firebrick")) + facet_wrap(vars(Site))

ggplot(data = difference_means) + 
  geom_histogram(aes(x = diff_se, fill = Treatment), position="stack", binwidth = 0.01)  +
  geom_vline( xintercept =  0)  +
  labs(title = "Difference in standard error of NDVI -- by Treatment",  x = "se NDVI (full plot) - se NDVI (tagged 5m transect)") +
  scale_fill_manual(values=c("CT" = ct_gray, "CT1" = ct_gray, "CT2"= ct_gray, "GH-CT" = ct_gray,
                              "N" = n_yellow, "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = p_blue,
                              "S" = "purple1", "L" = "purple2",
                              "NP" = lmat_colors[5],
                              "F0.5" = lmat_colors[1],
                              "F1" = lmat_colors[2],
                              "F2" = lmat_colors[3],
                              "F5" = lmat_colors[4],
                              "F10" = lmat_colors[5])) +
  facet_wrap(vars(Site))


# ## site summaries 
# 
# site_ndvi_all <- all_ndvi_means %>% 
#   group_by(Site, Treatment) %>% 
#   summarise_at(.vars = vars(mean), .funs = funs(mean,sd,se=sd(.)/sqrt(n()))) %>% 
#   mutate(measurement_area = "full_side")
# 
# site_ndvi_tagged <- tagged_means %>% 
#   group_by(Site, Treatment) %>% 
#   summarise_at(.vars = vars(mean), .funs = funs(mean,sd,se=sd(.)/sqrt(n()))) %>% 
#   mutate(measurement_area = "tagged_5m_transect")


############################ PLOTS 
```


### Boxplots by Site showing plot differences
```{r, echo=F, message=F}
## Boxplot
boxplot_data <- bind_rows(all_ndvi, tagged_ndvi) 

box_plot_het_function <- function(boxplot_data, check_site, facet_by_block = T, show_all_points = F) {
  boxplot_subset <- boxplot_data %>% 
    filter(Site == check_site) 
  
  plot <- ggplot(data = boxplot_subset, aes(x=Treatment, y=NDVI, color = measurement_area)) +
    geom_boxplot() +
    {if(show_all_points) geom_point(data = boxplot_subset, alpha = 0.5)} +
    {if(facet_by_block) facet_wrap(vars(Block))} +
    labs(title = check_site)

  return(plot)
}

box_plot_het_function(boxplot_data, "LMAT", facet_by_block = T, show_all_points =F)

box_plot_het_function(boxplot_data, "MAT", facet_by_block = T, show_all_points = F)

box_plot_het_function(boxplot_data, "HIST", facet_by_block = T, show_all_points = F)

box_plot_het_function(boxplot_data, "DHT", facet_by_block = T, show_all_points = F)

box_plot_het_function(boxplot_data, "MNAT", facet_by_block = T, show_all_points = F)

box_plot_het_function(boxplot_data, "NANT", facet_by_block = T, show_all_points = F)

box_plot_het_function(boxplot_data, "WSG", facet_by_block = T, show_all_points = F)


```

### Seasonal Trend 
Plot the seasonal trend when including data from whole plot. 


```{r, echo = F, message = F, fig.width = 12, fig.height=20}
#PLOT SUBSET -- SPECIFY SITE/DATE/ETC to ZOOM IN ON

check_trtmts <- c("CT", "CT1", "CT2", "F0.5", "F1", "F2", "F5", "F10", "N", "P", "NP")  


## PLOT OPTIONS
# Color Palettes 
n_yellow <- rgb(255, 192, 0, maxColorValue = 255)
p_blue <- rgb(46, 117, 182, maxColorValue = 255)
np_green <- rgb(112, 173, 71, maxColorValue = 255) #lmat_colors[5] 
ct_gray <- rgb(175, 171, 171, maxColorValue = 255)

lmat_colors <- c(rgb(226, 240, 217, maxColorValue = 255), rgb(169, 209, 142, maxColorValue = 255), rgb(112, 173, 71, maxColorValue = 255), rgb(84, 130, 53, maxColorValue = 255), rgb(56, 87, 35, maxColorValue = 255))


### DATA 
ndvi_plot <- ndvi_data %>% left_join(unispec_file_key) %>% 
  filter(Block %in% c("B1", "B2", "B3", "B4")) %>% 
  filter(Treatment %in% check_trtmts) %>% 
  # filter(!is.na(Treatment)) %>% 
  filter(!str_detect(Site, "LAB2|SHB2|MAT-SH")) %>% 
  mutate(Site = ifelse(Site %in% c("WSG1", "WSG23"), "WSG", Site))  %>% 
  mutate(Treatment = ifelse(Treatment %in% c("CT1", "CT2", "CT"), "CT", Treatment)) %>% 
  mutate(Treatment = factor(Treatment, levels = check_trtmts)) %>% 
  group_by(Site, Block, Treatment, Date) %>% 
  summarise_at(.vars = vars(NDVI), .funs = funs(mean,sd,se=sd(.)/sqrt(n()))) %>% 
  mutate(measurement_area = "full_side")

ndvi_plot_tagged <- ndvi_data %>% left_join(unispec_file_key) %>% 
  filter(Block %in% c("B1", "B2", "B3", "B4")) %>% 
  filter(Treatment %in% check_trtmts) %>% 
    filter(str_detect(ScanLocation, pattern = "[0-9]S")) %>% 
  filter(!str_detect(Site, "LAB2|SHB2|MAT-SH")) %>% 
    mutate(Site = ifelse(Site %in% c("WSG1", "WSG23"), "WSG", Site))  %>% 
  mutate(Treatment = ifelse(Treatment %in% c("CT1", "CT2", "CT"), "CT", Treatment)) %>% 
  mutate(Treatment = factor(Treatment, levels = check_trtmts)) %>% 
  group_by(Site, Block, Treatment, Date) %>% 
  summarise_at(.vars = vars(NDVI), .funs = funs(mean,sd,se=sd(.)/sqrt(n()))) %>% 
  mutate(measurement_area = "tagged_transect")
  

ndvi_comp <- bind_rows(ndvi_plot, ndvi_plot_tagged) 

(ndvi_ggplot <- ggplot(data = ndvi_comp, mapping = aes(x = Date, y = mean, color=Treatment)) +
  geom_line(aes(linetype=measurement_area)) + 
  geom_errorbar( aes(ymin = mean-se, ymax= mean + se, linetype=measurement_area), width=2) + 
  geom_point(aes(shape=measurement_area)) +
  facet_grid(Site ~ Block) + 
  theme_gray(base_size = 14) +
  scale_color_manual(values=c("CT" = ct_gray, "CT1" = ct_gray, "CT2"= ct_gray,
                              "N" = n_yellow, "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = p_blue,
                              "NP" = lmat_colors[5],
                              "F0.5" = lmat_colors[1],
                              "F1" = lmat_colors[2],
                              "F2" = lmat_colors[3],
                              "F5" = lmat_colors[4],
                              "F10" = lmat_colors[5])) +
  labs(y = "Mean NDVI") )

ndvi_ggplot 

```


## Drone Comparison
```{r}

band_defns <- tribble(
  ~definition, ~color, ~min, ~max,
  "ITEX", "red", 560, 600,
  "ITEX", "nir", 725, 1000,
  "MODIS", "red", 620, 670, 
  "MODIS", "nir", 841, 876,
  "MODIS", "blue", 459,479,
  "SKYE", "red", 620, 680,
  "SKYE", "nir", 830, 880,
  "SKYE", "blue", 455, 480,
  "Greenseeker", "red", 656, 656,
  "Greenseeker", "nir", 774, 774, 
  "RapidSCAN", "red", 670, 670,
  "RapidSCAN", "nir", 780, 780,
  "RapidSCAN", "red_edge", 730, 730, 
  "ToolikGIS_Drone_2018", "red", 640, 680,
  "ToolikGIS_Drone_2018", "nir", 820, 890,
  "ToolikGIS_MicaSense_2019", "blue", 455, 495,
  "ToolikGIS_MicaSense_2019", "green", 540, 580,
  "ToolikGIS_MicaSense_2019", "red", 658, 678,
  "ToolikGIS_MicaSense_2019", "red_edge", 707, 727,
  "ToolikGIS_MicaSense_2019", "nir", 800, 880,
  "ToolikEDC", "red", 560, 680,
  "ToolikEDC", "nir", 725, 1000
)
## Load relevant data
## 2019-07-15 : LMAT B3, B4

## Processed Data 
processed_dataframe <- read_rds(paste0(data_path, "2019_processed_spu_data.rds")) %>% 
  mutate(DateTime = lubridate::as_datetime(DateTime, tz="US/Alaska")) 

## Unispec Field Key
field_key <- paste0(data_path, "2019_unispec_field_key.csv")

unispec_field_key <- read_csv(field_key, skip_empty_rows = T, col_types = cols(
  Date = col_character(),
  Site = col_character(),
  Block = col_character(),
  Treatment = col_character(),
  ScanRep = col_double(),
  ScanLocation = col_character(),
  FileNum = col_double(),
  Weather = col_character(),
  Notes = col_character()
)) ## remove NA rows 

# Confirm spu_dataframe has same Site abbreviation as key
# > Standardize Site Names -----------------------------------------------------
# Check for different spelling of site names

unique(spu_dataframe$Site) %>% sort()
# spu_dataframe$Site <- recode (spu_dataframe$Site, !!!Site_Names, .default = spu_dataframe$Site)
unique(unispec_field_key$Site) %>% sort()

df <- full_join(spu_dataframe, unispec_field_key) %>% arrange(DateTime) %>% 
  mutate_at(.vars = vars(Site, Block, Treatment), .funs = factor) %>% 
  filter(Site %in% c("LMAT", "DHT"))

df_bands <- df %>% 
  mutate(Bands = map(Spectra, function(x) calculate_spectral_bands(x, band_defns = band_defns, instruments = c("ToolikGIS_MicaSense_2019"))))


drone_comp <- df_bands %>% 
  select(spu_filename, DateTime, Site, Date, Block, Treatment, ScanRep, ScanLocation, Notes, Weather, Bands) %>% unnest(Bands) %>% 
  filter(!is.na(Treatment), !str_detect(Treatment, "DARK|REF")) %>% 
  filter(str_detect(Date, "2019-07-15|2019-07-12")) %>% 
  mutate(AK_Time = lubridate::with_tz(DateTime, tzone = "US/Alaska"))


write_csv(drone_comp, paste0(data_path, "LTER_UnispecData_2019-07-12_2019-07-15_LMAT_DHT.csv"))

```

