---
title: "Unispec Notes & Summaries"
author: "Ruby An"
date: "5/4/2020"
output:
  html_document:
    df_print: paged
params:
  data_path: /home/ruby/ResearchProjects/Unispec/UnispecData/2019/
  function_file: unispec_protocol_functions.R
  session_date: '2019-06-08'
editor_options:
  chunk_output_type: console
---

```{r setup, include=F}
library(tidyverse)
library(lubridate)
library(rChoiceDialogs)
library(knitr)
library(plotly)

knitr::opts_chunk$set()

## use the params$data_path as starting folder. DATA PATH should be to UNISPEC DATA folder. The files should be read-in from there. This should match the One-Drive file format. 
data_path <- params$data_path

```

## Description

This document provides an overview of the Unispec Data collected by terrestrial RA Ruby An in 2017, 2018, 2019, and 2020. 

### 2017 Data Collection Notes
Unispec Data was collected approximately weekly. Unispec locations were changed by Jim and Ruby from historic locations to standardized spots, mainly outside of the destructive zones. See ---- file for map comparison. 

INSTRUMENT
PROTOCOL
LOCATIONS
DATA

#### Instrument Status: 
On -- the XXX foreoptic broke at the Shrub plots. Look for inconsistenies here with the rest of the season. 

#### Data File Descriptions

### 2018 Data Collection Notes
Unispec data was collected when weather permitted, weekly if possible. The summer was exceptionally rainy and vegetation was often wet, restricting the possible sampling times. At the end of the season, silver circular tags were nailed to the boardwalk at each unispec location (where line-segment perpendicular to the edge of the plot hit the boardwalk). This location was determined by measuring with a string and measuring tape along the plot edge between the wooden stakes marking each 5m segment of the plot. 

Drone data was collected in 2018 in conjunction with unispec measurements; however, this data

#### Instrument Status

turned out to be unusable due to a calibration error of the drone sensor. 

#### Data File Descriptions



### 2019 Data Collection Notes
Unispec data was collected approximately weekly. The MAT06 & MAT89 sites given highest priority, the historic and wet sedge least so. The SHB89 sites were only unispec'd once.  On ---- 15 scans were made per plot for comparison purposes. On ---- the default number of scans per plot was increased from 5 to 10 for all sites. It took a couple tries to land on "the most efficient" 10 spots to take the scans, so there is some variability over the course of the summer on where these "additional" measurements were made in each plot. 

#### Instrument Status

#### Data File Descriptions
The data collected in 2019 is saved in the following 4 files.  

  1. 2019_raw_spu_metadata.csv (instrument settings per .spu file)
  2. 2019_raw_spu_data.rds (raw spectra)
  3. 2019_processed_spu_data.rds (corrected spectra)
  4. 2019_index_data.csv (calculated vegetation indices)
  
Metadata correspondence key of scan (.spu file) to sample location. 

  5. 2019_unispec_coordinate_key.csv (no spu_filename column)
  6. 2019_unispec_file_key.csv (spu_filename column) 
  
Summary file for analysis

  7. 2019_unispec_data_summary.rds
  

```{r load_data, include=F}
## Load 2019 data from files

dir_year <- "2019"

## Raw Data
raw_filename <- paste0(data_path, dir_year, "_raw_spu_data.rds")
spu_spectra <- read_rds(raw_filename)
  
# Raw Metadata  
raw_keyname <- paste0(data_path, dir_year, "_raw_spu_metadata.csv")
spu_key <- read_csv(raw_keyname) %>% mutate(DateTime = with_tz(DateTime, "US/Alaska")) 

## Processed Data
processed_filename <- paste0(data_path, dir_year, "_processed_spu_data.rds")
processed_spectra <- read_rds(processed_filename) 

## Index Data
index_filename <- paste0(data_path, dir_year, "_index_data.csv")
index_data <- read_csv(index_filename) %>% mutate(DateTime = with_tz(DateTime, "US/Alaska"))


## File Key 
key_filename <- paste0(data_path, dir_year, "_unispec_file_key.csv")

unispec_file_key <- read_csv(key_filename, skip_empty_rows = T)

## Summary File 
df_analysis <- read_rds(paste0(data_path, dir_year,"_unispec_data_summary.rds"))
```

#### File Summaries
```{r}
## This chunk of code tallies the number of files in the raw, processed, and index data dataframes and displays the output in a table. 

## Raw Data summary

spu_dataframe <- full_join(spu_spectra, spu_key) %>% 
  mutate(DateTime = with_tz(DateTime, "US/Alaska")) %>% 
  
  # Add Date & ARC-LTER Site
  mutate(Date = lubridate::date(DateTime)) %>% 
  mutate(Site = str_extract(spu_filename, "([A-Z]{3,8}[0-9]*)(?=_)")) %>% # DOES NOT WORK FOR "MAT-SH" measurements
  mutate(Site = ifelse(str_detect(spu_filename, "MAT-SH"), "MAT-SH", Site))

raw_files <- spu_dataframe %>%   group_by(Date, Site) %>%
  summarize(Raw_Files = n_distinct(spu_filename))

## Processed Data summary
processed_files <- left_join(processed_spectra, spu_key) %>%
  
  # Add Date & ARC-LTER Site
  mutate(DateTime = with_tz(DateTime, "US/Alaska")) %>% 
  mutate(Date = lubridate::date(DateTime)) %>% 
  mutate(Site = str_extract(spu_filename, "([A-Z]{3,8}[0-9]*)(?=_)")) %>% # DOES NOT WORK FOR "MAT-SH" measurements
  mutate(Site = ifelse(str_detect(spu_filename, "MAT-SH"), "MAT-SH", Site)) %>% 
  
  ## Summarize
  group_by(Date, Site) %>%
  summarize(Processed_Files = n_distinct(spu_filename))

## Index Data summary
index_files <- left_join(index_data, spu_key) %>%   
  
  # Add Date & ARC-LTER Site
  mutate(DateTime = with_tz(DateTime, "US/Alaska")) %>% 
  mutate(Date = lubridate::date(DateTime)) %>% 
  mutate(Site = str_extract(spu_filename, "([A-Z]{3,8}[0-9]*)(?=_)")) %>% # DOES NOT WORK FOR "MAT-SH" measurements
  mutate(Site = ifelse(str_detect(spu_filename, "MAT-SH"), "MAT-SH", Site)) %>% 
  
  ## Summarize
  group_by(Date, Site) %>%
  summarize(Index_Files = n_distinct(spu_filename))

## DISPLAY
kable(full_join(raw_files, processed_files) %>% full_join(index_files) %>% arrange(Date, Site))


```



Code chunks:

  1. Correct Spectra
  2. Calculate Indices
  3. Format data for analyses / storage
  4. Format data for visualization
  

#### For Shiny
```{r, echo=F}

# Subset of Data for Shiny Visualization
(
  shiny_dataframe <- index_data %>%
    select(-BandDefinition) %>%
    spread(Index, Value) %>%
    left_join(unispec_file_key) %>%

    # Check for tag
    #filter(str_detect(ScanLocation, pattern = "[0-9]S")) %>%

    # Format for shiny
    rename(Replicate = ScanRep) %>%
    select(DateTime, Date, Site, Block, Treatment, Replicate, FileNum, NDVI, EVI, EVI2, Plot.y) %>%
    mutate(Year = lubridate::year(DateTime),
           # Date = lubridate::ymd(Date),
           DOY = lubridate::yday(DateTime)) %>%
    rename(Time = DateTime) %>%
    mutate(Block = as.numeric(str_extract(Block, "\\d"))) %>%
    mutate(Replicate = as.character(Replicate)) %>%

    # Standardize Site names from 2019 version to 2020 onwards
    mutate(Site = ifelse(Site %in% c("WSG1", "WSG23"), "WSG89", Site))  %>%
    mutate(Site = ifelse(Site %in% c("DHT"), "DHT89", Site)) %>% 
    mutate(Site = ifelse(Site %in% c("MAT"), "MAT89", Site)) %>% 
    mutate(Site = ifelse(Site %in% c("LMAT"), "MAT06", Site)) %>% 
    mutate(Site = ifelse(Site %in% c("HIST"), "DHT89", Site)) %>% 
    mutate(Site = ifelse(Site %in% c("SHB2"), "SHB89", Site)) %>% 
    mutate(Site = ifelse(Site %in% c("MNAT"), "MNT97", Site)) %>% 
    mutate(Site = ifelse(Site %in% c("NANT"), "MNN97", Site))

)

plot_data <- shiny_dataframe %>% 
  filter(Site %in% c("MAT06"),Treatment %in% c("CT", "CT1", "CT2", "F10", "NP")) %>% 
  filter(Replicate %in% c(1:10)) %>% 
  filter(Plot.y > 9 && Plot.y < 17) %>% # original 5 measurement 
  mutate(Spot = paste0(Treatment, "_", Plot.y))

plot <- ggplot(data = plot_data, aes(x = DOY, 
                                     y = NDVI, 
                                     color = Treatment,
                                     text = paste(
                          "Location: ", Plot.y, "\n",
                          sep = ""))) + 
  geom_point(aes(alpha = 0.2)) + 
  geom_line(aes(group=c(Spot), alpha=0.2)) +
  geom_smooth(aes(group=Treatment)) +
  facet_grid(Site ~ Block)

plot
ggplotly(plot)

```

### 2020 Data Collection Notes 
Unispec data collected once at the following sites: 

- Sites: MAT06, MAT89, DHT89, MNT97, MNN97, WSG89
- Dates: 2020-07-16 to 2020-07-19
- Protocol: 10/plot using Unispec-DC and new field keys (SHB89: RapidScan 5/plot)
Data Processing: “UnispecR” git-hub repository, piloted shiny-app 

#### Instrument Status
Unispec performed fine. The downward facing foreoptic seemed to be breaking a bit at the joint to the metal. However, checking the Darkscans it didn't seem like any light was leaking in. 

#### Data File Descriptions




