---
title: "Compare Multispec vs. R Processing Pipelines"
author: "Ruby An"
date: "March 19, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Required Packages
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE)

## Required Packages
require("tidyverse")
require("knitr")
require("lubridate")
source("unispec_record_functions.R") # file loads required functions

## Vectors
WSG <- c("WSG1", "WSG23", "WSG2")
SHB <- c("SHB1", "SHB2")

## Standardize SiteNames
Site_Names <- list(DH ="HTH", DHT = "HTH", LHTH = "HTH", HTHB = "HTH", HTHPC = "HTH",HIS="HIST", 
                   LOF = "LMAT",  LOFB = "LMAT", LNB = "LMAT", LOFRB ="LMAT",
                   MATB="MAT", MATSL= "MAT", MATBK = "MAT", 
                   MANTB ="MNAT",MNATB ="MNAT", NAMTB = "MNAT", 
                   NMNT = "NANT", NANTB ="NANT", JULNB ="NANT",NMNTB ="NANT",
                   LSHB= "SHB", SHBB = "SHB", SHRBB = "SHB", SHRB = "SHB", 
                   LWSG = "WSG", WSGB = "WSG", WS ="WSG", WSB = "WSG", WSDB = "WSG",
                   WSG1 = "WSG", WSG23 = "WSG", WSG2 = "WSG",
                   SHB1 = "SHB", SHB2 = "SHB")
```

# Read Multispec Data
```{r}

data_path <- "UnispecData/2018"

# 1. Read in File Key --------------------------------------------------------
# This gives you the correspondence between each .spu file and the 
# site / block / plot / treatment / measurement at which it was taken. 

# Find all file keys 
key_files <- list.files(path = data_path, pattern = "*_key.csv*", full.names = T, recursive = T)

# Read in filekeys 
key_list <- tibble(keyname = key_files) %>% # create dataframe
  mutate(key_contents = map(keyname, function(x) read_key_file(x)))


keys <- unnest(key_list) # %>% 
  # mutate(Site = replace(Site, Site %in% WSG, "WSG")) %>% 
  # mutate(Site = replace(Site, Site %in% SHB, "SHB")) 


# 2. Multispec File Lists ----------------------------------------------------

# Create file lists (SITE-YEAR-MONTH-DAY-multispecstate.csv)
multispec_files <- list.files(path = data_path, pattern = "raw|correct", full.names=T, recursive = T)

bad_files <- c("UnispecData/2018/multispec/Processed/2018-07-03_DHT_correct.csv",
               "UnispecData/2018/multispec/Processed/2018-07-03_NANT_B3_correct.csv",
               "UnispecData/2018/multispec/Processed/2018-07-03_NANT_correct.csv",
               "UnispecData/2018/multispec/Processed/2018-07-03_WSG1_correct.csv",
               "UnispecData/2018/multispec/Raw/2018-07-03_DHT_raw.csv",
               "UnispecData/2018/multispec/Raw/2018-07-03_NANT_B3_raw.csv",
               "UnispecData/2018/multispec/Raw/2018-07-03_NANT_raw.csv",
               "UnispecData/2018/multispec/Raw/2018-07-03_WSG1_raw.csv",
               "UnispecData/2018/multispec/Raw/2018-08-01_DHT_raw.csv")

good_files <- multispec_files[!(multispec_files %in% bad_files)]
# 3. Read in data ---------------------------------------------------------

data_list <- tibble(filename = good_files) %>% # create dataframe
  mutate(file_contents = map(filename, function(x) read_multispec_file(x)))

data <- unnest(data_list) #%>% 
  #mutate(Site = replace(Site, Site %in% WSG, "WSG")) %>% 
  #mutate(Site = replace(Site, Site %in% SHB, "SHB")) 

## Join to File Key to get block, treatment, measurement
keys_data <- inner_join(keys, data)

multispec_data  <- keys_data %>% 
  mutate(multispec_filename = filename) %>% 
  mutate(Status = ifelse(Type == "correct", "multispec_corrected", "multispec_raw")) %>% 
  select(-c(filename, keyname, Type)) 
  #filter(!(Treatment %in% c("REF", "DARK"))) %>% # Exclude the reference & dark

multispec_dataframe <- "UnispecData/2018_multispec_dataframe.rds"
write_rds(multispec_data, path = multispec_dataframe)

```

# Load Multispec & Rscript Data

```{r unispec_data}

# LOAD DATA ---------------------------------------------------------------
df_rscript <- read_rds("UnispecData/2018_unispec_dataframe_clean.rds") %>% 
  # Remove columns 
  select(-raw_indices, -corrected_indices, -DateTime, -Integration_ms, -ref_integration_ms) %>% 
  # Rename columns 
  rename(raw = raw_spectra,
         corrected = corrected_spectra,
         rscript_ref_filenames = ref_filenames) %>% 
  gather(key = Status, value = Spectra, raw, corrected)  %>% 
  # formate filenames
  mutate(spu_filename = str_remove(spu_filename, "UnispecData/2018/spu/")) %>% 
  mutate(rscript_ref_filenames = str_remove_all(rscript_ref_filenames, "UnispecData/2018/spu/[0-9]+(-[0-9]+)+/"))


df_multispec <- read_rds("UnispecData/2018_multispec_dataframe.rds") %>% 
  filter(!multispec_filename == "UnispecData/2018/multispec/Processed/2018-07-22_NANT_B1B2_correct.csv") %>% #duplicate multispec file, different references used
  rename(multispec_ref_filenames = ref_files) %>% 
  separate(Status, sep = "_", into = c("ProcessType", "Status"), fill = "right") %>% 
  select(-ProcessType) %>% 
  unnest(Spectra) %>% 
  rename(Multispec_Reflectance = Reflectance)

# Standardize Names
df_rscript$Site <- recode (df_rscript$Site, !!!Site_Names, .default = df_rscript$Site)
unique(df_rscript$Site)
df_multispec$Site <- recode (df_multispec$Site, !!!Site_Names, .default = df_multispec$Site)
unique(df_multispec$Site)
```

## Interpolate Rscript & Join 
```{r}
# Interpolate raw & rscript-corrected data
df_interp <- df_rscript %>% 
  filter(Treatment != "DARK") %>% 
  mutate(Spectra = map(Spectra, function(x) interpolate_spectra(x))) %>% 
  unnest(Spectra) %>% 
  rename(Interp_Reflectance = Reflectance)

# Join Multispec & Rscript-Interpolated data
df_comp <- inner_join(df_multispec, df_interp) %>% 
  # Absolute Difference between multispec vs rscript interpolated values 
  mutate(Difference_Reflectance = Multispec_Reflectance - Interp_Reflectance) %>% 
   # Percentage difference (of reflectance value )between multispec vs rscript interpolated
  mutate(Difference_Percent = 2*Difference_Reflectance/(Multispec_Reflectance + Interp_Reflectance))

# Summary of Multispec vs Rscript Spectra differences
df_comp %>% summary()

```

## Save comparison data
```{r}
## Nest data for filtering 
df_comp_nest <- df_comp %>% 
  nest(Wavelength, Multispec_Reflectance, Interp_Reflectance, Difference_Reflectance, Difference_Percent, .key = Spectra)

## Save data
filename <- "UnispecData/2018_multispec_vs_rscript.rds"
write_rds(df_comp_nest, filename)
```


# Spectra Comparison

## Compare White References used
```{r}

## Check for different references
df_comp_refs <- df_comp_nest %>% # filter(abs(Difference_Reflectance) > 0.01) %>% 
  filter(Status == "corrected") %>% 
  select(Date, Site, Block, multispec_ref_filenames, rscript_ref_filenames) %>% 
  arrange() %>% 
  distinct() %>% 
  mutate(refs_same = map2(multispec_ref_filenames, rscript_ref_filenames, check_same_references)) %>% 
  mutate(refs_same = unlist(refs_same))

## {Dates, Sites, Blocks} where multispec & rscript refs differ
df_refs_diff <- df_comp_refs %>% filter(!refs_same) %>% select(-refs_same) 
```

## Check for Spectra Differences
```{r}
## Check for differences in Reflectance Spectra
df_diff <- df_comp_nest %>% anti_join(df_refs_diff) %>% filter(Status == "corrected") %>% 
  unnest(Spectra) %>% 
  filter(abs(Difference_Reflectance) > 0.01) %>%  # check for greater than 1% difference in reflectance
  ## factor for summaries 
  mutate_at(vars(Date, Site, Block, Treatment, Replicate, spu_filename, multispec_filename), funs(factor)) %>% 
  arrange(Difference_Percent)

## List spu_files with multispec vs rscript spectra differences 
df_diff_files <- df_diff %>% group_by(spu_filename) %>% 
  summarize(count = n(), avg_wavelength = mean(Wavelength), avg_difference = mean(Difference_Reflectance), percent_difference = mean(Difference_Percent*100))

## Summary of egregiously different data
df_diff_files
df_diff %>% summary()

hist(df_diff$Difference_Percent)
```

## Plot Spectra Differences 
```{r}

## Plot differences 
first_file <- 1
num_files_to_plot <- 3

df_diff_plot <- df_comp_nest %>% inner_join(df_diff_files %>% select(spu_filename)) %>% 
  arrange(spu_filename) %>% 
  slice(first_file:(first_file+num_files_to_plot)) %>% 
  unnest(Spectra) %>% 
  gather(key = "ProcessType", value = "Value", Multispec_Reflectance, Interp_Reflectance, Difference_Reflectance, Difference_Percent) %>%
  #filter(ProcessType != "Difference_Percent") %>% 
  mutate(Unit = ifelse(str_detect(ProcessType, "Difference"), "Difference", "Reflectance"))
  # separate(ProcessType, sep = "_", into = c("ProcessType", "Unit"))

df_rscript_plot <- df_rscript %>% 
  inner_join(df_diff_files %>% select(spu_filename)) %>% 
  arrange(spu_filename) %>% 
  slice(first_file:(first_file+num_files_to_plot)) %>% 
  unnest(Spectra) %>% 
  gather(key = "ProcessType", value = "Value", ChA, ChB, Reflectance) %>% 
  mutate(Unit = ifelse(str_detect(ProcessType, "Ch"), "Channel", "Reflectance")) 


## PLOT SELECTED DATA 
color_set <- c("ChA" = "skyblue", "ChB" = "lightgreen", "Difference_Reflectance" = "red", 
               "Interp_Reflectance" = "purple", "Multispec_Reflectance" = "orange", "Reflectance" = "grey")
ggplot(data = df_diff_plot, mapping = aes(x = Wavelength, y = Value)) +
  geom_point(data = df_rscript_plot, aes(color=ProcessType, alpha = Status), size = 0.75) +
  geom_line(aes(color=ProcessType, alpha = Status)) +
  scale_alpha_manual(values = c(1, 0.3)) +
  #scale_color_manual(values = color_set) +
  facet_grid(Unit ~ spu_filename, scales = "free") 

ggplot(data)
```


Main conclusion: Multispec does something beyond solely interpolating the raw values. Linear interpolation in R and multispec raw interpolated files show deviations in spectra. 

However, multispec corrected and r-corrected then interpolated values are (in most but not all) cases very similar [INSERT STATISTIC HERE]. 

Differences usually occurred at [????] Wavelengths. 

# NDVI Comparison
```{r}
# NDVI Comparison ----------------------------------------------------------
### No evidence for difference between NDVI between processing types. 
### Lines for "multispec" vs "rscript" plot right over one another.
### Do actual stats on this? repeated measures ANOVA or dependent t-test? 

## PLOT SELECTION over DATES
sites <-c("MAT") #
blocks <- c("B1") # c("B1", "B2", "B3", "B4")
trtmts <- c("CT1") # c("CT", "NP", "N", "P")
measures <- c("1", "2")

ndvi_types <- df_types %>% 
  filter(Site %in% sites) %>% 
  filter(Block %in% blocks) %>% 
  filter(Treatment %in% trtmts) %>% 
  calculate_ndvi_process()

ndvi_plot <- ndvi_types

ggplot(data = ndvi_plot, mapping = aes(x = Date, y = ndvi, color = Type, linetype = ProcessType)) +
  geom_point() +
  geom_line() + 
  facet_grid( Measurement ~ Type)


## PLOT SELECTION for Individual Date 
### change the following vectors to explore all data over dates
dates <- df_types %>% select(Date) %>% unique() %>% slice(1:n()) %>% c()
dates <- dates[[1]]
sites <-c("MAT") #
blocks <- c("B1") # c("B1", "B2", "B3", "B4")
trtmts <- c("CT1") # c("CT", "NP", "N", "P")
measures <- c("1", "2", "3", "4", "5")

ndvi_types <- df_types %>% 
  filter(Date %in% dates) %>% 
  filter(Site %in% sites) %>% 
  filter(Block %in% blocks) %>% 
  filter(Treatment %in% trtmts) %>% 
  calculate_ndvi_process()

ndvi_plot <- ndvi_types

ggplot(data = ndvi_plot, mapping = aes(x = ProcessType, y=ndvi, fill=Type)) +
  geom_bar(stat="identity") +
  facet_grid( Measurement ~ Type)
```

