---
title: "Process LTER Unispec Data"
author: "Ruby An"
date: "December 17, 2018"
output:
  html_notebook:
    number_sections: yes
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

This R Markdown document visualizes and does stats on unispec data collected from 2010-2018. 

# Required Packages
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE)

## Required Packages
require("tidyverse")
require("knitr")
require("lubridate")
source("unispec_record_functions.R") # file loads required functions
```

For now, I am assuming R corrected spectra and multispec corrected spectra are the same. I am only using the multispec corrected spectra, as this is available for all years. 


# Load Data
Load historic unispec data (clean). 

```{r unispec_data}
setwd("/home/ruby/UnispecR/UnispecRecord/")
df_2014 <- read_rds("UnispecData/2014_unispec_dataframe_clean.rds") %>% 
  rename(raw_spectra = spu_spectra, 
         corrected_spectra = multispec_spectra)
df_2015 <- read_rds("UnispecData/2015_unispec_dataframe_clean.rds") %>% 
  rename(corrected_spectra = multispec_spectra)
df_2016 <- read_rds("UnispecData/2016_unispec_dataframe_clean.rds") %>% 
    rename(raw_spectra = spu_spectra,
           corrected_spectra = multispec_spectra)
df_2017 <- read_rds("UnispecData/2017_unispec_dataframe_clean.rds") 
df_2018 <- read_rds("UnispecData/2018_unispec_dataframe_clean.rds")

df_2019 <- read_rds("UnispecData/2019_unispec_dataframe_clean.rds")


## Transform for problem check
df <- df_2017 %>% 
  rename(Spectra = raw_spectra) %>% 
  mutate(xlsx_filename = NA) 

unispec_key_fix <- read_key_file("UnispecData/2017/2017_unispec_key.csv")

## FOR NICOLE: Save 2017 NDVI for excel 
ndvi_for_excel <- df_2017 %>% unnest(corrected_indices) %>% filter(Index == "NDVI") %>% rename(NDVI = Value) %>% select(-Index)

write_csv(ndvi_for_excel, "UnispecData/2017_ndvi_for_excel.csv")

day_averages <- ndvi_for_excel %>% select(Date, Site, Block, Treatment, Replicate, NDVI) %>% 
  group_by(Date, Site, Block, Treatment) %>% 
  summarize(mean_ndvi = mean(NDVI))

write_csv(day_averages, "/home/ruby/2017_ndvi_plot_averages.csv")


## MISSING REP 1 for 2017 data LMAT 
ndvi_for_excel %>% filter(Date == "2017-07-31", Site == "LMAT", Block == "B1") %>% 
  print(n=100)






```

# Save Data for Shiny Visualization
```{r, df_shiny} 

## 2014
## Load new data for shiny
df_2014 <- read_rds("UnispecRecord/UnispecData/2014_unispec_data.rds") %>% 
    filter(! multispec_spectra %>% map(is.null) %>% map_lgl(any)) %>% 
  select(-Spectra) %>% 
  rename(Spectra = multispec_spectra) %>% 
  ## Calculate NDVI
  mutate(Indices = map(Spectra, function(x) calculate_indices(x, band_defns = band_defns, instrument = "MODIS", indices = c("NDVI", "EVI", "EVI2"))))

indices_2014 <- df_2014
write_rds(indices_2014, "UnispecRecord/UnispecData/2014_index_data.rds")

unispec_problem_key_2014 <- read_csv("UnispecRecord/UnispecData/2014_unispec_key_problems.csv", skip_empty_rows = T, col_types = cols(ref_problems = col_character())) %>% 
  mutate_at(.funs = factor, .vars = vars("Site", "Block", "Treatment", "Replicate", "key_fix", "ref_problems")) %>% 
  left_join(read_csv("UnispecRecord/UnispecData/2014_raw_spu_key.csv") %>% select(spu_filename, DateTime))

unispec_data_2014 <- unispec_problem_key_2014 %>% right_join(df_2014)

indices_2014 <- unispec_data_2014 %>%  
  mutate(Year = as.integer(lubridate::year(DateTime))) %>% 
  mutate(Date = lubridate::date(DateTime)) %>%
  mutate(DOY = as.integer(lubridate::yday(Date))) %>% 
  mutate(Block = as.integer(str_extract(Block, pattern = "[0-9]"))) %>% 
  mutate(Replicate = as.character(Replicate)) %>% 
  rename(Time = DateTime) %>% 
  unnest(Indices) %>% spread(Index, Value) %>% 
  select(Year, Date, DOY, Time, Site, Block, Treatment, Replicate, FileNum, NDVI, EVI, EVI2)

indices_2014

## 2015
## Calculate Index Data from processed dataframe
df_2015 <- read_rds("UnispecRecord/UnispecData/2015_processed_xlsx_data.rds")

unispec_problem_key_2015 <- read_csv("UnispecRecord/UnispecData/2015_unispec_key_problems.csv", skip_empty_rows = T, col_types = cols(ref_problems = col_character())) %>% 
  mutate_at(.funs = factor, .vars = vars("Site", "Block", "Treatment", "Replicate", "key_fix", "ref_problems")) %>% 
  left_join(read_csv("UnispecRecord/UnispecData/2015_raw_spu_key.csv") %>% select(spu_filename, DateTime))

unispec_data_2015 <- unispec_problem_key_2015 %>% 
  filter(!is.na(Treatment)) %>% 
  filter(!file_problem) %>% left_join(df_2015) %>%
  filter(! multispec_spectra %>% map(is.null) %>% map_lgl(any)) %>% 
  rename(Spectra = multispec_spectra)

indices_2015 <- unispec_data_2015 %>% 
  ## Calculate NDVI
  mutate(Indices = map(Spectra, function(x) calculate_indices(x, band_defns = band_defns, instrument = "MODIS", indices = c("NDVI", "EVI", "EVI2"))))

write_rds(path = "UnispecRecord/UnispecData/2015_index_data.rds", indices_2015)

indices_2015 <- indices_2015 %>% 
  mutate(Year = as.integer(lubridate::year(Date))) %>% 
  mutate(Date = lubridate::date(Date)) %>%
  mutate(DOY = as.integer(lubridate::yday(Date))) %>% 
  mutate(Block = as.integer(str_extract(Block, pattern = "[0-9]"))) %>% 
  mutate(Replicate = as.character(Replicate)) %>% 
  rename(Time = DateTime) %>% 
  unnest(Indices) %>% spread(Index, Value) %>% 
  select(Year, Date, DOY, Time, Site, Block, Treatment, Replicate, FileNum, NDVI, EVI, EVI2)

## 2016
unispec_problem_key_2016 <- read_csv("UnispecRecord/UnispecData/2016_unispec_key_problems.csv", skip_empty_rows = T, col_types = cols(ref_problems = col_character())) %>% 
  mutate_at(.funs = factor, .vars = vars("Site", "Block", "Treatment", "Replicate", "key_fix", "ref_problems")) %>% 
  left_join(read_csv("UnispecRecord/UnispecData/2016_raw_spu_key.csv") %>% select(spu_filename, DateTime))


indices_2016 <- read_rds("UnispecRecord/UnispecData/2016_index_data.rds") %>% 
  left_join(unispec_problem_key_2016) %>% 
  filter(!file_problem)  %>% 
  mutate(Year = as.integer(lubridate::year(Date))) %>% 
  mutate(Date = lubridate::date(Date)) %>%
  mutate(DOY = as.integer(lubridate::yday(Date))) %>% 
  mutate(Block = as.integer(str_extract(Block, pattern = "[0-9]"))) %>% 
  mutate(Replicate = as.character(Replicate)) %>% 
  rename(Time = DateTime) %>% 
  unnest(Indices) %>% 
  filter(spu_filename != "jul21lof00006.spu")%>% spread(Index, Value) %>% 
  select(Year, Date, DOY, Time, Site, Block, Treatment, Replicate, FileNum, NDVI, EVI, EVI2)

## 2017 
indices_2017 <- read_rds("UnispecRecord/UnispecData/2017_index_data.rds") %>%
  mutate(Year = as.integer(lubridate::year(Date))) %>% 
  mutate(Date = lubridate::date(Date)) %>%
  mutate(DOY = as.integer(lubridate::yday(Date))) %>% 
  mutate(Block = as.integer(str_extract(Block, pattern = "[0-9]"))) %>% 
  mutate(Replicate = as.character(Replicate)) %>% 
  rename(Time = DateTime) %>% 
  unnest(Indices) %>% spread(Index, Value) %>% 
  select(Year, Date, DOY, Time, Site, Block, Treatment, Replicate, FileNum, NDVI, EVI, EVI2)

## 2018 
indices_2018 <- read_rds("UnispecRecord/UnispecData/2018_unispec_dataframe_clean.rds") %>%   mutate(Year = as.integer(lubridate::year(Date))) %>% 
  mutate(Date = lubridate::date(Date)) %>%
  mutate(DOY = as.integer(lubridate::yday(Date))) %>% 
  mutate(Block = as.integer(str_extract(Block, pattern = "[0-9]"))) %>% 
  mutate(Replicate = as.character(Replicate)) %>% 
  rename(Time = DateTime) %>% unnest(corrected_indices) %>% 
  spread(Index, Value) %>% 
  select(Year, Date, DOY, Time, Site, Block, Treatment, Replicate, FileNum, NDVI, EVI, EVI2) 

## 2019 
indices_2019 <- read_rds("UnispecRecord/UnispecData/2019/2019_index_data.rds") %>%
  mutate(Year = as.integer(lubridate::year(Date))) %>% 
  mutate(Date = lubridate::date(Date)) %>%
  mutate(DOY = as.integer(lubridate::yday(Date))) %>% 
  mutate(Block = as.integer(str_extract(Block, pattern = "[0-9]"))) %>% 
  mutate(Replicate = as.character(Replicate)) %>% 
  rename(Time = DateTime) %>% 
  unnest(Indices) %>% spread(Index, Value) %>% 
  select(Year, Date, DOY, Time, Site, Block, Treatment, Replicate, FileNum, NDVI, EVI, EVI2) 

## All Data 
shiny_data <- read_rds("UnispecRecord/Visualizations/shiny_unispec_cleaned/unispec_indices_summary_2014-2018.rds")

shiny_data <- bind_rows(indices_2014, indices_2015, indices_2016, indices_2017, indices_2018, indices_2019) %>% 
  filter(!is.na(DOY))

## Standardize Names 
Site_Names <- list(DHT = "HTH", DH ="HTH", LHTH = "HTH", HTHB = "HTH", HTHPC = "HTH", HST = "HIST", HIS="HIST", 
                   LOF = "LMAT",  LOFB = "LMAT", LNB = "LMAT", LOFRB ="LMAT",
                   MATB="MAT", MATSL= "MAT", MATBK = "MAT", 
                   MANTB ="MNAT",MNATB ="MNAT", NAMTB = "MNAT", 
                   NMNT = "NANT", NANTB ="NANT", JULNB ="NANT",NMNTB ="NANT",
                   LSHB= "SHB", SHBB = "SHB", SHRBB = "SHB", SHRB = "SHB", 
                   LWSG = "WSG", WSGB = "WSG", WS ="WSG", WSB = "WSG", WSDB = "WSG", TEST = "LAB2", WSG1 = "WSG", WSG23 = "WSG", SHB1 = "SHB", SHB2 = "SHB")

# Check for different spelling of site names
unique(shiny_data$Site)
shiny_data$Site <- recode (shiny_data$Site, !!!Site_Names, .default = shiny_data$Site)

unique(shiny_data$Site)

write_rds(shiny_data, "UnispecRecord/Visualizations/shiny_unispec_cleaned/indices_2014-2019.rds")


index_data <- read_rds("UnispecRecord/Visualizations/shiny_unispec_cleaned/indices_2014-2019.rds")
```

# Check Spectral Data

## Plot Spectra: Site vs. Block
For the following code, make sure you select only one date per site. The 5 measurements per plot are averaged as a line and one standard deviation above and below shaded. Interactively edit the "PLOT SELECTION" vectors in the code chunk below to investigate specific data.


```{r plot_spectra_blocks, dependson="unispec_data"}
# Recall df_table for list of dates/sites
df_table <- df_clean %>% 
  select(Date, Site) %>% 
  distinct() %>% 
  group_by(Date) %>% 
  do(Sites = t(.[-1])) %>% 
  mutate(Sites = paste( unlist(Sites), collapse=','))
kable(df_table) 

#PLOT SUBSET -- SPECIFY SITE/DATE/ETC to ZOOM IN ON
check_dates <- ymd("2016-07-02", "2016-06-23")
check_sites <- c("MAT", "LMAT", "WSG")
check_blocks <- c("B1", "B2", "B3", "B4")
check_trtmts <- c("NP", CT) 

# Data Comparison Format
df_block <- df_clean %>% 
  filter(Date %in% check_dates) %>% 
  filter(Site %in% check_sites) %>% 
  filter(Block %in% check_blocks) %>% 
  filter(Treatment %in% check_trtmts) %>% 
  unnest(spu_spectra) %>% 
  group_by(Site, Block, Treatment, Date, Wavelength) %>% 
  summarize(
    avg_reflect = mean(Reflectance),
    max_ref = max(Reflectance),
    min_ref = min(Reflectance),
    sd_reflect = sd(Reflectance),
    n = n()
    ) 

ggplot(data = df_block, mapping = aes(x = Wavelength, y = avg_reflect)) +
  geom_line(aes(color=Treatment)) + 
  geom_ribbon(aes(ymin=avg_reflect-sd_reflect, ymax=avg_reflect+sd_reflect, fill=Treatment), alpha=0.25) +
  facet_grid(Date ~ Site + Block ) + 
  scale_color_manual(values=c("CT" = "forestgreen", "CT1"="darkolivegreen2", "CT2"="darkolivegreen3",
                              "N" = "dodgerblue", "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = "red2",
                              "NP" = pur_pal[5],
                              "F0.5" = pur_pal[1],
                              "F1" = pur_pal[2],
                              "F2" = pur_pal[3],
                              "F5" = pur_pal[4],
                              "F10" = pur_pal[5])) + 
  scale_fill_manual(values=c("CT" = "forestgreen", "CT1"="darkolivegreen2", "CT2"="darkolivegreen3",
                              "N" = "dodgerblue", "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = "red2",
                              "NP" = pur_pal[5],
                              "F0.5" = pur_pal[1],
                              "F1" = pur_pal[2],
                              "F2" = pur_pal[3],
                              "F5" = pur_pal[4],
                              "F10" = pur_pal[5]))

```


 
## Plot Spectra: Site vs. Date
Average plot averages by block. NOTE: Ask Laura what the correct error propogation is here.

```{r plot_spectra_dates, echo=F}
#PLOT SUBSET -- SPECIFY SITE/DATE/ETC to ZOOM IN ON
check_dates <- ymd("2016-06-18", "2016-06-23")
check_sites <- c("MAT", "LMAT")
check_blocks <- c("B1", "B2", "B3", "B4")
check_trtmts <- c("NP", CT, "N", "P") 

# Data Comparison Format
df_dates <- df_clean %>% 
  filter(Date %in% check_dates) %>% 
  filter(Site %in% check_sites) %>% 
  filter(Block %in% check_blocks) %>% 
  filter(Treatment %in% check_trtmts) %>%
  unnest(spu_spectra) %>% 
  group_by(Site, Block, Treatment, Date, Wavelength) %>% # average measurements by plot
  summarize(
    avg_reflect = mean(Reflectance),
    max_ref = max(Reflectance),
    min_ref = min(Reflectance),
    sd_reflect = sd(Reflectance)
    )  %>% 
  group_by(Site, Treatment, Date, Wavelength) %>% # average plots by block
  summarize(
    block_avg_reflect = mean(avg_reflect),
    max_ref = max(avg_reflect),
    min_ref = min(avg_reflect),
    sd_reflect = sd(avg_reflect)
    ) 

ggplot(data = df_dates, mapping = aes(x = Wavelength, y = block_avg_reflect)) +
  geom_line(aes(color=Treatment)) + 
  geom_ribbon(aes(ymin=block_avg_reflect-sd_reflect, ymax=block_avg_reflect+sd_reflect, fill=Treatment), alpha=0.25) +
  facet_grid(Site ~ Date) + 
  scale_color_manual(values=c("CT" = "forestgreen", "CT1"="darkolivegreen2", "CT2"="darkolivegreen3",
                              "N" = "dodgerblue", "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = "red2",
                              "NP" = pur_pal[5],
                              "F0.5" = pur_pal[1],
                              "F1" = pur_pal[2],
                              "F2" = pur_pal[3],
                              "F5" = pur_pal[4],
                              "F10" = pur_pal[5])) + 
  scale_fill_manual(values=c("CT" = "forestgreen", "CT1"="darkolivegreen2", "CT2"="darkolivegreen3",
                              "N" = "dodgerblue", "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = "red2",
                              "NP" = pur_pal[5],
                              "F0.5" = pur_pal[1],
                              "F1" = pur_pal[2],
                              "F2" = pur_pal[3],
                              "F5" = pur_pal[4],
                              "F10" = pur_pal[5]))

```


# Check Vegetation Indices

Currently, this only works for NDVI, EVI, and EVI2 as I haven't worked out spectral interpolation yet and the other indices need reflectance at a specific value (not a range). 

## Extract NDVI data
```{r convert_ndvi}
## Only NDVI data
ndvi_data <- df_clean %>%  # select just NDVI 
  unnest(multispec_indices) %>% 
  filter(Index == "NDVI(MODIS)") %>% 
  rename(NDVI = Value) 
```

## Plot NDVI: Site vs. Block
```{r ndvi, echo=F}
#PLOT SUBSET -- SPECIFY SITE/DATE/ETC to ZOOM IN ON
check_dates <- ndvi_data %>% select(Date) %>% unique() %>%  pull()#ymd("2017-06-08", "2017-06-15", "2017-06-26")
check_sites <- c("MAT")
check_blocks <- c("B1", "B2", "B3", "B4")
check_trtmts <- c("NP", "F10", CT) 

# ndvi_types <- df_tidy %>% # to calculate NDVI 
#   filter(Date %in% check_dates) %>% 
#   filter(Site %in% check_sites) %>% 
#   filter(Block %in% check_blocks) %>% 
#   filter(Treatment %in% check_trtmts) %>%
#   calculate_index(indices = "NDVI") # function in "unispec_functions.R"

ndvi_plot <- ndvi_data %>% 
  filter(Date %in% check_dates) %>%
  filter(Site %in% check_sites) %>%
  filter(Block %in% check_blocks) %>%
  filter(Treatment %in% check_trtmts) %>%
  group_by(Site, Block, Treatment, Date) %>% 
  summarise(
    avg_ndvi = mean(NDVI),
    sd_ndvi = sd(NDVI)
  ) %>% 
  group_by(Site,Treatment,Date) %>% 
  summarise(
    avg_ndvi = mean(avg_ndvi),
    sd_ndvi = sd(sd_ndvi)
  )

ndvi_subset <- ndvi_data %>% 
  filter(Date %in% check_dates)%>%
  filter(Site %in% check_sites) %>%
  filter(Block %in% check_blocks)%>%
  filter(Treatment %in% check_trtmts) %>% 
  group_by(Block)

ggplot(data = ndvi_subset, mapping = aes(x = Date, y = NDVI, color = Treatment)) +
  geom_point(aes(shape=Block), position = "jitter") + 
  geom_smooth(aes(fill=Treatment), method = lm, formula = y ~ splines::bs(x, 3)) + 
  facet_wrap(vars(Site)) + 
  scale_color_manual(values=c("CT" = "forestgreen", "CT1"="darkolivegreen2", "CT2"="darkolivegreen3",
                              "N" = "dodgerblue", "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = "red2",
                              "NP" = pur_pal[5],
                              "F0.5" = pur_pal[1],
                              "F1" = pur_pal[2],
                              "F2" = pur_pal[3],
                              "F5" = pur_pal[4],
                              "F10" = pur_pal[5])) +
    scale_fill_manual(values=c("CT" = "forestgreen", "CT1"="darkolivegreen2", "CT2"="darkolivegreen3",
                              "N" = "dodgerblue", "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = "red2",
                              "NP" = pur_pal[5],
                              "F0.5" = pur_pal[1],
                              "F1" = pur_pal[2],
                              "F2" = pur_pal[3],
                              "F5" = pur_pal[4],
                              "F10" = pur_pal[5])) 
  
  
plot_errorbar <- ggplot(data = ndvi_plot, mapping = aes(x = Date, y = avg_ndvi, color=Treatment, fill=Treatment)) +
  geom_point() +
  geom_line() + 
  geom_errorbar(aes(ymin = avg_ndvi-sd_ndvi, ymax= avg_ndvi + sd_ndvi), width=2) + 
  facet_wrap(vars(Site)) + 
  scale_color_manual(values=c("CT" = "forestgreen", "CT1"="darkolivegreen2", "CT2"="darkolivegreen3",
                              "N" = "dodgerblue", "NO3" = "skyblue", "NH4" = "deepskyblue",
                              "P" = "red2",
                              "NP" = pur_pal[5],
                              "F0.5" = pur_pal[1],
                              "F1" = pur_pal[2],
                              "F2" = pur_pal[3],
                              "F5" = pur_pal[4],
                              "F10" = pur_pal[5])) 
```


# Compare Raw|Correct NDVI

## Calculate Indices from Spectra
```{r df_ndvi}
# build tibble of #band_defns: instrument, color, min, max
band_defns <- tribble(
  ~definition, ~color, ~min, ~max,
  "ITEX", "red", 560, 600,
  "ITEX", "nir", 725, 1000,
  "MODIS", "red", 620, 670, 
  "MODIS", "nir", 841, 876,
  "MODIS", "blue", 459,479,
  "SKYE", "red", 620, 680,
  "SKYE", "nir", 830, 880,
  "SKYE", "blue", 455, 480,
  "ToolikGIS_Drone_2018", "red", 640, 680,
  "ToolikGIS_Drone_2018", "nir", 820, 890
)

ndvi_from_spectra <- clean_unispec_dataframe %>% 
  filter(!str_detect(Treatment, "REF")) %>% 
  mutate(spu_indices = map(spu_spectra, function(x) calculate_indices(x, band_defns = band_defns, instrument = "MODIS", indices = "NDVI")))

df_ndvi  <- ndvi_from_spectra %>% 
  # mutate(multispec_indices = map(multispec_indices, function(x) filter(x, Index=="NDVI(MODIS)"))) %>% 
  unnest(spu_indices) %>% 
  mutate(Index = str_c(Index, "_raw")) %>% 
  unnest(multispec_indices) %>% 
  mutate(Index1 = str_c(Index1, "_corrected")) %>% 
  gather(key = temp, value = Index, Index, Index1) %>% # combine indices into one column 
  gather(key = temp, value = Value, Value, Value1) %>% 
  select(-temp) 

```

## Plot NDVI w/Errorbars 
```{r ndvi_comp}

#PLOT SUBSET -- SPECIFY SITE/DATE/ETC to ZOOM IN ON
check_dates <- df_ndvi %>% select(Date) %>% unique() %>%  pull()#ymd("2017-06-08", "2017-06-15", "2017-06-26")
check_sites <- c("MAT", "LMAT", "HIST")
check_blocks <- c("B1", "B2", "B3", "B4")
check_trtmts <- c(CT, "NP") 


df_ndvi_subset <- df_ndvi %>% 
  filter(Date %in% check_dates)%>%
  filter(Site %in% check_sites) %>%
  filter(Block %in% check_blocks)%>%
  filter(Treatment %in% check_trtmts) %>% 
  filter(str_detect(Index, "NDVI"))
  
df_ndvi_summary <- df_ndvi_subset %>% 
  ## optional: group CT plots
  mutate(Treatment = replace(Treatment, Treatment %in% CT, "CT")) %>%  
  ## average data
  group_by(Site, Block, Treatment, Date, Index) %>% 
  summarise(
    avg_ndvi = mean(Value),
    sd_ndvi = sd(Value)
  ) %>% 
  group_by(Site,Treatment,Date,Index) %>% 
  summarise(
    avg_ndvi = mean(avg_ndvi),
    sd_ndvi = sd(sd_ndvi)
  ) 

ggplot(data = df_ndvi_summary, mapping = aes(x = Date, y = avg_ndvi, color=Index)) +
  geom_point() +
  geom_line() + 
  geom_errorbar(aes(ymin = avg_ndvi-sd_ndvi, ymax= avg_ndvi + sd_ndvi)) +
  facet_grid(Treatment ~ Site) 

```

