---
title: "Process Unispec Files collected on One Day"
author: "Ruby An"
date: "2021-06-24"
output: html_document
editor_options: 
  
  chunk_output_type: console
params:
  session_date: 2020-07-16
  data_path: /home/ruby/UnispecData/2020/raw_data/2020-07-16 # select folder where spu files reside
---

## Introduction
This file is fo  processing unispec data in a user friendly way. 

## Required Packages & Functions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F)

## Required Packages
library("tidyverse")
library("knitr")
library(rChoiceDialogs)
library(lubridate)

## R FUNCTIONS for processing spu data
read_spu_file_metadata <- function(filename, info = "short") {
  # DESCRIPTION: Reads first 9 text lines in .spu files 
  # INPUT: .spu file -- collected using PPSystems UnispecDC
  #         info = "short", the default returns only the spu_filename, DateTime, FileNum, Integration time
  #                "long", returns all the info in the 9 lines of header in .spu files
  # OUTPUT: dataframe with 15 columns on instrument status, scan settings, max/min value

    # Extract info from the file itself, reading metadata from first 9 lines. Create a dataframe
    text <- read_lines(filename, n_max=9)
    
    # Line 1: Extract the file name in the spu file as a check. Some file names have spaces 
    spu_filename <- str_replace(text[1],".*[\\\\]([A-z0-9.]\\s*)","\\1") %>% # extract filename
      str_replace("\"","") # removes trailing quote at end of line 
    FileNum <- str_extract(spu_filename, "\\d{5}") %>% as.numeric() # from 5 digits in filename
    
    # Line 2: 
    Remarks <- str_split(text[2], pattern = " ")[[1]] # split by "space"
    Type <- str_split(Remarks[7], pattern = "=")[[1]][2] # extract relevant part
    ScanType <- ifelse(grepl("DARKscan",Type, fixed=T), "DARKscan", # format
                       ifelse(grepl("Datascan,DC",Type, fixed=T), "Throwawayscan", 
                              Type))
    DarkscanID <- str_extract(text[2], "Dark=.+.spu")
    Remarks <- text[2]
    
    # Line 3: 
    DateTime <-  lubridate::mdy_hms(text[3], tz="America/Anchorage")
    
    # Line 4: Limits -- range of spectra measured
    Limits <- str_extract(text[4], "\\d+.\\d.+\\d") 
    
    # Line 5: 
    Temperature <- as.numeric(strsplit(strsplit(text[5], split = " ")[[1]][4], split="=")[[1]][2])
    Battery <- str_extract(text[5], "BattV=\\d+.\\d+")
    Aux <- str_extract(text[5], "A\\d=.+\\d")
    
    # Line 6-9: 
    Minimum <- str_extract(text[6], "\\d+.+\\d") # Wavelength, ChB min
    Minimum_wavelength <- str_split(Minimum, boundary("word"))[[1]][1] # Wavelength
    Minimum_value <- str_split(Minimum, boundary("word"))[[1]][2] # ChB AD
    
    Maximum <- str_extract(text[7], "\\d+.+\\d") # Wavelength, ChB max
    Maximum_wavelength <- str_split(Maximum, boundary("word"))[[1]][1]
    Maximum_value <- str_split(Maximum, boundary("word"))[[1]][2]

    Integration <- as.numeric(strsplit(text[8], split = " ")[[1]][3])
    NumberScans <- str_extract(text[9], "\\d+")
    
    # Truncated Filename - use as SCANID to join to other dataframes
    spu_filename <- unlist(str_split(filename, pattern = "/")) %>% last()
    
    # Metadata 
    metadata <- tibble(spu_filename, DateTime, FileNum, ScanType, Integration, NumberScans,Minimum_wavelength,Minimum_value, Maximum_wavelength, Maximum_value, Limits, Temperature, Battery, Aux, DarkscanID, Remarks)
    
    if(info == "short") {
      metadata <- metadata %>% select(spu_filename, DateTime, FileNum, Integration)
    }

  # Print filenames while reading 
  #print(spu_filename) # use for error checking
  
  return(metadata)
}


read_spu_file_spectra <- function(filename) {
  # DESCRIPTION: For a generic .spu file regardless of name, extract spectral data
  # INPUT: Unispec-DC .spu file
  # OUTPUT: dataframe of spectral data with 3 columns Wavelength, ChB, ChA
  
  # Read spectral intensity data into dataframe
  data <- read.table(file = filename, skip = 9, col.names = c("Wavelength", "ChB", "ChA"))
  
  print(filename)
  
  return(data)
}

# Color band definitios for calculate_indices function
band_defns <- tribble(
  ~definition, ~color, ~min, ~max,
  "ITEX", "red", 560, 600,
  "ITEX", "nir", 725, 1000,
  "MODIS", "red", 620, 670, 
  "MODIS", "nir", 841, 876,
  "MODIS", "blue", 459,479,
  "SKYE", "red", 620, 680,
  "SKYE", "nir", 830, 880,
  "SKYE", "blue", 455, 480,
  "ToolikGIS_Drone_2018", "red", 640, 680,
  "ToolikGIS_Drone_2018", "nir", 820, 890,
  "ToolikGIS_MicaSense_2019", "blue", 455, 495,
  "ToolikGIS_MicaSense_2019", "green", 540, 580,
  "ToolikGIS_MicaSense_2019", "red", 658, 678,
  "ToolikGIS_MicaSense_2019", "red_edge", 707, 727,
  "ToolikGIS_MicaSense_2019", "near_ir", 800, 880,
  "ToolikEDC", "red", 560, 680,
  "ToolikEDC", "nir", 725, 1000
)

calculate_indices <- function(spectra, band_defns, instrument = "MODIS", indices = "NDVI") {
  # Calculates NDVI, EVI, and EVI2 from dataframe including Wavelength : Spectra 
  ## inputs: spectra - Wavelength, Reflectance columns
  ##         band_defns : wavelengths definining colors 
  ##         instrument : e.g. MODIS, SKYE, ITEX
  ##         indicies   : the index to return 
  ## output: Index - name of vegetation index
  ##         BandDefinition - name of "instrument" or spectral band definition used
  ##         Value - value of index, with the band definition used. 
  
  bands <- band_defns %>% 
    filter(definition == instrument) 
  
  blue <- bands %>% filter(color=="blue") %>% select(min, max) %>% as.numeric()
  nir <- bands %>% filter(color=="nir") %>% select(min, max) %>% as.numeric()
  red <- bands %>% filter(color=="red") %>% select(min, max) %>% as.numeric()
  
  spectra_bands <- spectra %>% 
    mutate(color = ifelse(Wavelength >= blue[1] & Wavelength <= blue[2], "blue",
                          ifelse(Wavelength >= red[1] & Wavelength <= red[2], "red",
                                 ifelse(Wavelength >= nir[1] & Wavelength <= nir[2], "nir",
                                        "other")))) %>% 
    group_by(color) %>% 
    summarize(Reflectance = mean(Reflectance))
  
  index_data <- spectra_bands %>%
    spread(color, Reflectance) %>% 
    mutate(NDVI = (nir-red)/(nir+red),
           EVI = 2.5*((nir-red)/(nir+6*red-7.5*blue + 1)),
           EVI2 = 2.5*((nir-red)/(nir+2.4*red + 1))) %>% 
    select_at(indices) %>% 
    gather(Index, Value, everything()) %>% 
    
    # Add Spectral Band Definition convention
    mutate(BandDefinition = instrument) %>% 
    select(Index, BandDefinition, Value)
  
  return(index_data) 
}


## SANITY CHECK FUNCTIONS

check_scan_times <- function(raw_data) {
  
  timedata <- raw_data %>% 
    select(Site, DateTime, FileNum) %>% 
    distinct(DateTime, .keep_all = T)
  
  # Calculate time between scans 
  timedata$diff <- timedata$DateTime - lag(timedata$DateTime)
  
  meta_timedata <- left_join(timedata, field_keys)
  
  time_check <- meta_timedata %>% select(Site, DateTime, Block, Treatment, Replicate, FileNum, diff, everything()) %>% ungroup()
  
  return(time_check)
}

check_for_missing_spectra <- function(spu_dataframe) {
  missing <- spu_dataframe %>% filter(Spectra %>% map(is.null) %>% map_lgl(any))

  return(missing)
}

```

## Read in data
  
  * data_path is year folder where the session files are stored. The structure is year/spu/yyyy-mm-dd  where yyyy-mm-dd is sample date. Should be on OneDrive.
  
  * Field Key file with header "Date,Site,Block,Treatment,Replicate,Location,FileNum,Notes,Weather".  File name convention is year_unispec_field_key.csv

### Read .SPU files 
  
```{r, include = F}
# Choose folder with .spu files
data_path <- rchoose.dir(caption = "Selecte directory containing .spu files")

## Make a character vector of all .spu raw files in current session date folder
# Check if directory exist and use it as the working directory
spu_files <- list.files(path = data_path, pattern = ".spu$", full.names = T, recursive=T)

# Read metadata text lines (9) from the spu files
spu_filedata <- map_dfr(spu_files, read_spu_file_metadata)

# Add Metadata columns:
#   spu_filename_full: variable with the full path filename 
#   Site: column from spu_filename
#   FileNum: column from spu_filename 
#   ScanType: column identifying scan type (Dark, Throwaway, Datascan)
spu_metadata <- spu_filedata %>% 
  mutate(spu_filename_full = spu_files) %>% #
  mutate(Site = toupper(str_extract(spu_filename, "[A-Za-z]{3}[0-9]{1,2}(?=_)"))) %>% # get string that is 3 letters, 2 numbers before _ 
  mutate(FileNum = as.integer(str_extract(spu_filename, "\\d{5}"))) 

# Read spectra from .spu files > add to metadata
spu_dataframe <- spu_metadata %>% 
  mutate(Spectra=map(spu_filename_full, function(x) read_spu_file_spectra(x))) %>% 
  mutate(Date = date(DateTime))
  # mutate(Time = DateTime-floor_date(DateTime, unit = "day")) 

```

### Read Unispec File Key
```{r}

# Choose key files
key_files <- list.files(path = data_path, pattern = "*unispec_key.csv$", full.names = T, recursive=T)

# Read key files 
key_files <- rchoose.files(caption = "Choose Unispec File Key", multi = T)

field_keys <- key_files %>%  purrr::map(function(file_name) read_csv(file_name)) %>% 
  reduce(rbind) %>% 
  mutate(Date = date(params$session_date)) # Add date


# Join by SITE, DATE, FILENUM
raw_data <- left_join(spu_dataframe, field_keys) %>% arrange(DateTime) %>% 
  mutate_at(.vars = vars(Site, Block, Treatment), .funs = factor)
```


### QUALITY CHECK: Unispec Key & .spu file Correspondance  

```{r missing, warnings=F}
# This chunk prints a bunch of checks at the end. Read through it to check that nothing is amiss. 

### MISSING: Identify any measurement locations that are missing corresponding raw spu files. By checking the overlap between "field_keys" and "spu_dataframe", this code lists (a) missing raw .spu files OR (b) extra .spu files that aren't accounted for in the unispec field keys. 


## Identify any missing files (or extra files)
missing_spu_files <- anti_join(field_keys, spu_dataframe)

## Is anything missing from the field keys? 
#### Look for NA's indicating missing info
missing_info <- raw_data %>% 
  filter(is.na(spu_filename) | # spu_filename
         is.na(Site) | # site
         is.na(Block) & !str_detect(Treatment, "REF|DARK|VEG") | # Block NA's should always be REFS or EXTRA
         is.na(Replicate) & !str_detect(Treatment, "REF|DARK|THROWAWAY"), # Check for replicate NA's that aren't REF
         Treatment != "EXTRA|VEG|REF") %>% # don't care about EXTRA, VEG, or REF scans
  select(spu_filename, FileNum, Site, Block, Treatment, Replicate, everything()) # reorder columns to see output better

### Identify any files that aren't listed in the field key
extra_files <- anti_join(spu_dataframe, field_keys) %>% pull(spu_filename)


### MISLABELING: Check for human error in recording location correspondence between metadata and .spu file numbers. 

## Check File Number Pattern -- The number of files per block should be multiples of 5, unless REF or NA. This code identifies any plots that are not multiples of 5.
df_filenum_count <- field_keys %>% group_by(Site, Date, Block, Treatment) %>% 
    summarize(Num_Files = n()) %>% # count the number of files
    filter(Num_Files %% 5 != 0) %>%   # files per plot
    
    # Remove scans that aren't of plot locations (DARK, REF, THROWAY)
    filter(Treatment != "THROWAWAY") %>% # filter out scantypes that don't correspond to plot measurement locations
    filter(Treatment != "DARK") %>% 
    filter(! "REF" %in% Treatment) 


# Print output 
## Confirm the number of spu files 
print(paste0("Total number of .spu files: ", nrow(spu_dataframe)))
print(paste0("Total number of scans documented: ", nrow(field_keys)))
## Missing things
print("QUALITY CHECKS: all answers should be zero if things ran well")
print(paste0("Missing .spu files: ", nrow(missing_spu_files)))
print(paste0("Extra .spu files: ", length(unprocessed_files)))
print(paste0("Scan Locations missing info: ", nrow(missing_info)))
## Mislabeling
print(paste0("Plots with an abnormal number of scans (not a clean multiple of 5): ", nrow(df_filenum_count)))


```


## Correct data with REFerence files

White references correct for instrument & cable irregularities. Multiplying by the correction factor (ChA/ChB) smooths out the spectra. There are typically 5 reference measurements per *Date* / *Site*. If multiple file numbers are listed, the correction factors are averaged. 

Based on the original field notebook key and the following quality checks, choose reference files by entering the appropriate file numbers in **`r field_keys`** for the rows where the column *Treatment* = **REF**. 

Make sure to rerun the **Read Unispec Keys** if you change REF file choices. The following code plots your chosen references for a visual check.

### Choose & Check References
```{r ref_choice_list}

## NOTE: DELETED CODE that allowed for:
# (a) using references not taken at the site during the same session, e.g. if you took references at MAT06 and then used those same files for MAT89. 
# (b) code that allows you to match different ref spectra to different blocks, e.g. in the case that the direction the unispec is facing matters. This code ignores that possibilty (some evidence for this in 2019 checks) for now. 

ref_files <- raw_data %>% filter(Treatment == "REF") 

ref_data <- ref_files %>% 
  
  ### Unnest Spectra & calculate correction factor
  unnest(Spectra) %>% 
  
  mutate(correction_factor = ChA/ChB) %>% 

  ### Group repeated REF measurements based on your plot set-up (choose Block or NOT)
  group_by(Date,Site,Integration,Wavelength) 


## Average 5 chosen ref measurements per DATE/SITE & calculate mean correction factors

correction_factors <- ref_data %>% 
  summarize(correction_factor = mean(ChA/ChB), 
            ref_filenames = str_c(spu_filename,collapse = ", "))


## QUALITY CHECK
## Table per Site of Reference Files 
(ref_filenames_table <- ref_data %>% group_by(Date, Site, Integration) %>% 
  summarize(Files = n_distinct(spu_filename), ref_filenames = str_c(spu_filename,collapse = ", ")) )

## Build Plot all reference data files
(ref_plot <- ggplot(ref_data, aes(x = Wavelength, y = correction_factor, group=spu_filename)) + 
  theme(legend.position="left") + 
  geom_line(aes(color=factor(Integration))))

```

### Apply References
Rerun the **Join Data & keys** sections above to update the `raw_data` dataframe. Apply your chosen references to actual spectral data to create the tidy dataframe `processed_data` containing corrected sepectral reflectance values.

```{r apply_refs, echo=FALSE}

## Join DATA with REFS

raw_data_with_refs <- raw_data %>% 
  filter(!str_detect(Treatment, "REF|DARK|THROWAWAY")) %>% # remove ref & dark scans
  unnest(Spectra) %>% 
  filter(Wavelength > 400, Wavelength < 1000) %>% # remove edge wavelengths, instrument unreliable at extremes
  
  # join with ref data
  left_join(correction_factors) %>%
  mutate(raw_reflectance = ChB/ChA) %>% # the raw reflectance
  mutate(corrected_reflectance = raw_reflectance*correction_factor) # corrected reflectance

## Corrected Reflectances 
processed_data <- raw_data_with_refs %>% 
  nest(processed_spectra = c(Wavelength, ChB, ChA, correction_factor, raw_reflectance, corrected_reflectance))
```


##### Plot Check
```{r}

(processed_data %>% 
  slice(1:10) %>% 
  unnest(processed_spectra) %>% 
  gather(key = Status, value = Reflectance, raw_reflectance, corrected_reflectance) %>% 
  
  # VIZ
  ggplot(mapping = aes(x = Wavelength, y = Reflectance )) + 
  geom_line(aes(color = spu_filename, linetype = Status)) + 
    facet_wrap(vars(Date, Site, Block, Treatment, FileNum)))

```


## ? INTERPOLATION


## SAVE DATA
```{r}
processed_spu_data <- processed_data %>%
  select(spu_filename, DateTime, ref_filenames, processed_spectra) %>% unnest(processed_spectra) %>%
  select(spu_filename, DateTime, ref_filenames, Wavelength, ChB, ChA, raw_reflectance, CorrectionFactor, corrected_reflectance) %>%
  nest(Spectra = c(Wavelength, ChB, ChA, raw_reflectance, CorrectionFactor, corrected_reflectance))

ref_spu_data <- ref_data %>% unnest(Spectra) %>%
  select(spu_filename, DateTime, Wavelength, ChB, ChA) %>%
  mutate(Reflectance = ChB/ChA) %>% 
  mutate(CorrectionFactor = 1/Reflectance, corrected_reflectance = NA) %>%
  filter(Wavelength > 400, Wavelength < 1000) %>%
  rename(raw_reflectance = Reflectance) %>%
  nest(Spectra = c(Wavelength, ChB, ChA, raw_reflectance, CorrectionFactor, corrected_reflectance))

processed_spectra <- bind_rows(processed_spu_data, ref_spu_data)

## Save
processed_spu_filename <- paste0(data_path, "/", session_year, "_processed_spu_data.rds")

write_rds(df_processed, processed_spu_filename)
```


## Calculate INDICES

Currently, this only works for NDVI, EVI, and EVI2 as I haven't worked out spectral interpolation yet and the other indices need reflectance at a specific value (not a range).

### Calculate Indices
```{r}

df_processed %>%  filter( Spectra %>% map(is.null) %>% map_lgl(any)) # remove files w/out spu_spectra

df_indices <- df %>% select(-Spectra) %>% #dataframe with metadata
  filter(!is.na(Treatment)) %>%
  filter(!str_detect(Treatment, "REF|DARK")) %>%
  inner_join(df_processed) %>% # corrected reflectances with 
  ## Format for calculuating indices
  unnest(Spectra) %>%
  select(-ChB, -ChA, -raw_reflectance, -CorrectionFactor) %>%
  rename(Reflectance = corrected_reflectance) %>%
  nest(Spectra = c(Wavelength, Reflectance)) %>%
  ## Calculate NDVI
  mutate(Indices = map(Spectra, function(x) calculate_indices(x, band_defns = band_defns, instrument = "MODIS", indices = c("NDVI", "EVI", "EVI2"))))

index_filename <- paste0(data_path, "/", params$session_date, "_index_data.rds")
write_rds(df_indices, index_filename)


## convert to SHINY ready format


```


## Rename Raw Files for Long-Term Storage 
Rename files ("SITE_FileNum.spu") to include parent directory (Date) in filename: "YYYY-MM-DD_SITE_FileNum.spu". Copy the following code and use in the command line (haven't figured out how to run in RStudio yet... some kind of error that "file doesn't exist"). Adjust so that year (date) structure is correct. 

Do this before processing, so that the filenames match the long-term filenames. 
```{bash}

cd /UnispecData/2020  # go to directry with session "date folders"
#Set $SESSION = session date somehow 

#mkdir raw_data

# RUN THIS TO RENAME FILES AND MOVE THEM INTO .spu folders 
#!/bin/bash
for date in 2020-07-19; do
  pushd "$date"
  for file in *; do
    cp "$file" ../"${date}_${file}"
  done
  popd
done

  mv *.spu raw_spectra_files/ #make sure the "raw_data" folder exists first! :) 

```

### Visualize with other years

### Check across years

## CHECKLIST/SUMMARY REPORT



