---
title: "R Notebook"
output: html_notebook
---
Collection of code snippets used to clean and quality check unispec data. 

## CHOOSE FILES MANUALLY 
```{r}
## OR CHOOSE MANUALLY 
# Check if directory exist and use it as the working directory
if (!dir.exists(data_path) || !grepl(session_year,data_path)) {
  data_path <- rchoose.dir(caption = "Select Unispec Data directory")}
# Find unispec_key files 
key_files <- rchoose.files(default = data_path, caption = "Select unispec key file")
```


### Check compatability of Field Key with data files 
```{r, echo = F, warning=F}

# Confirm spu_dataframe has same Site abbreviation as key
if(!(unique(spu_dataframe$Site) %in% unique(field_keys$Site) %>% all())) {
  print("ERROR! SITE NAME in spu_files and spu_fieldkey differ! Check file names and file key template.")
}
u
# Date inconsistencies
if(spu_dataframe %>% pull(Date) %>% unique() != params$session_date){
  print("ERROR! Date Discrepancy")
  print(spu_dataframe %>% pull(Date) %>% unique())
}

```


## Data Summary 
```{r data_summary} 
## Data summary
spu_dataframe %>%
  mutate(Date = floor_date(DateTime, unit="day")) %>% 
  group_by(Date, Site) %>% 
  summarize(Files = n_distinct(spu_filename)) %>% 
  kable(caption = "Raw .spu files read-into spu_dataframe")

## Print Summary of files 
df %>% group_by(Date, Site, Block) %>%  
  filter(!is.na(Treatment)) %>% 
  summarize(Treatments = str_c(str_replace_na(sort(unique(Treatment))), collapse = ","), Num_Files = n()) %>% 
  kable(caption ="Files listed in field key")

```




## QAQC

### Missing files 
List missing raw .spu files and those that are unprocessed (raw .spu file does not appear in summary .xlsx). Missing raw files. 
```{r missing, warnings=F}
### MISSING: Find all processed data that is missing corresponding raw spu files 
missing_spu_data <- anti_join(field_keys, spu_dataframe)

if(nrow(missing_spu_data) > 0){
  missing_spu_data 
}

(unprocessed_files <- anti_join(spu_dataframe, field_keys) %>% pull(spu_filename))
```


### Mis-Labeling 
Check for human error in recording correspondence between metadata and file numbers. 

#### Discrepancy Checks
Check for possible mislabeling. 

```{r df_check, dependson="df", eval= T, echo=F, warnings=F}

## General NA test
df_na <- df %>% filter(ScanType == "Datascan") %>% 
  filter(is.na(spu_filename) | # spu_filename
           is.na(Site) | # site
           is.na(Block) & !str_detect(Treatment, "REF|DARK|VEG") | # Block NA's should always be REFS or EXTRA
           is.na(Replicate) & !str_detect(Treatment, "REF"), # Check for replicate NA's that aren't REF
         Treatment != "EXTRA|VEG|REF") # don't care about EXTRA 


## Check for large filenumbers : Inconsistent File Number reading due to number at end of site names
df_lgfn <- df %>%
  filter(FileNum > 600) %>%  # Sites w/numbers at the end mis-read in as part of file number. All > 100000: Unispec-DC measurements only go up to 5 digits.
  select(Date, Site, Block, Treatment, FileNum, spu_filename)


## Check File Number Pattern -- pulls out treatments that are not multiples of 5
#### Num Files per block should be multiples of 5, unless REF or NA
df_filenum_count <- field_keys %>% group_by(Site, Date, Block) %>% 
  filter(Treatment != "THROWAWAY") %>% 
  filter(Treatment != "DARK") %>% 
  filter(Treatment != "REF") %>% 
  summarize(Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), Num_Files = n()) %>% 
  filter(Num_Files %% 5 != 0)  # files per plot

```


Missing labels: 
`r if(nrow(df_na)>0){kable(df_na)}`
Large filenumbers: 
`r if(nrow(df_lgfn)>0){kable(df_lgfn)}`

An abnormal numbers of files per plot: 
`r if(nrow(df_filenum_count)>0){kable(df_filenum_count)}`

#### Duplicates 
Identify duplicates in field key or spu_files. Fix by editing unispec_key_fix.csv or deleting duplicate files. Only duplicates should be REF files used for other sites. 

```{r duplicates}

### Check for duplicate entries in key
duplicate_entries <- field_keys %>% group_by(Date, Site, Block, Treatment, Replicate, FileNum) %>%
  filter(n()>1) %>%
  arrange(Date, Site, FileNum) 

if(nrow(duplicate_entries)>0) {
  duplicate_entries %>% kable()
}

### Check raw spu_files have no duplicates

duplicates <- spu_dataframe %>% group_by(DateTime, FileNum) %>%
  filter(n()>1) %>%
  arrange(Date, Site, FileNum)

## File List 
duplicate_files <- duplicates$spu_filename
  
if (nrow(spu_dataframe) != length(unique(spu_dataframe$DateTime))) {
    ## Check for duplicate spu files 
  duplicates %>%   kable()
}


```

### Instrument Check:

#### Max'd Out Spectra (> 65000 AD)
List files that maxed out, in the wavelengths used to calculate MODIS NDVI.
```{r maxed, dependson="df"}

maxed_data <- df %>% inner_join(spu_dataframe %>% select(spu_filename)) %>% # restrict to those w/.spu files
  unnest(Spectra) %>% 
  filter(ChA > 65000 | ChB > 65000) %>% 
  group_by(spu_filename) %>% 
  summarize(maxed_number = n(), maxed_wavelengths = str_c(min(Wavelength), " - ", max(Wavelength), collapse = ", "))  

# Print MAXED files 
if (nrow(maxed_data) > 0) {
  df %>% select(spu_filename, Site, Date, Block, Treatment, Replicate) %>% 
     inner_join(maxed_data)
  
} else { print("no maxed files")}

# Maxed file list 
maxed_files <- maxed_data$spu_filename
```


#### Dim'd Out Spectra
Primarily Darkscans should show up. 
```{r dim, dependson=df, warnings=F}
dim_data <- df %>% inner_join(spu_dataframe %>% select(spu_filename)) %>% 
  #filter(!ScanType %in% c("Throwawayscan", "DARKscan")) %>% 
  unnest(Spectra) %>% 
  group_by(spu_filename) %>% 
  summarize(ChA_max = max(ChA)) %>% 
  filter(ChA_max < 20000) %>% 
  left_join(df) 

# Print DIM files  
if (nrow(dim_data) > 0) {
  dim_data %>% select(spu_filename, Site, Date, Block, Treatment, Replicate, ChA_max) %>% 
    kable()
  
} else { print("no dim files")}

# File List
dim_files <- dim_data$spu_filename
```


#### Time Check / Editable Data Table
Times between files, check for Mislabeling, etc.
```{r time_check}
# Select columns to Check
timedata <- df %>% 
  select(Site, Date, DateTime, FileNum, Integration, ScanType) %>% 
  distinct(DateTime, .keep_all = T)

timedata$diff <- timedata$DateTime - lag(timedata$DateTime)

meta_timedata <- left_join(timedata, field_keys)

time_check <- meta_timedata %>% select(Site, Date, DateTime, Block, Treatment, Replicate, FileNum, diff, everything()) %>% ungroup()

# Examine dataframe
time_check %>% select(Date, Site, Block, Treatment, FileNum, diff, Integration) %>% print(n=450)
```


#### Zero'd Spectra
```{r zerod}
## zero'd data with restricted wavelengths 
zero_data <- df %>% inner_join(spu_dataframe %>% select(spu_filename)) %>% unnest(Spectra) %>% 
  filter(ChA == 0) %>%  ## This is for all wavelengths, not just 400-1000nm
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  group_by(spu_filename) %>% 
  summarize(Zeros = n()) %>% 
  left_join(df)

# Print ZEROED files 
if (nrow(zero_data_narrowed) > 0) {
  zero_data_narrowed %>% select(spu_filename, Site, Date, Block, Treatment, Replicate, Zeros)
  
} else { print("no zero'd files")}
  
```


### Raw Spectra Plot Check
```{r check_refs}


df %>% 
  # arrange for 
  unnest(Spectra) %>% 
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  mutate(Reflectance = ChB/ChA) %>% 
  gather(key = Channel, value = Intensity, ChB, ChA) %>% 
  gather(key = ref_part, value = Reflectance_Intensity, Intensity, Reflectance) %>% 
  
  # Viz
  ggplot(mapping = aes(x = Wavelength, y = Reflectance_Intensity, group = FileNum)) +
  geom_line(aes(color = Treatment)) +
  facet_grid(ref_part ~ FileNum, scales="free") 


```

### Reference mistakes 
```{r}
ref_mistakes <- ref_data_mistakes %>% distinct(spu_filename) %>% pull()



## Build Plot all reference mistakes
ref_data_mistakes <- ref_data_all %>%
  filter(CorrectionFactor > 5) %>% 
  distinct(spu_filename) %>% 
  select(spu_filename) %>% 
  left_join(ref_data_all)

(ref_plot_mistakes <- ggplot(ref_data_mistakes, aes(x = Wavelength, y = CorrectionFactor)) +
  geom_line(aes(color=spu_filename)))

## PLOTS
ref_plot_all + ggtitle("ALL REFERENCES")
# ref_plot_mistakes + ggtitle("REF Mistakes") +
#   scale_y_continuous(limits = c(0, NA))

```


#### Reflectance >1 
```{r absurd}

## Corrected Spectra
absurd_data <- df_corrected %>% 
  filter(!str_detect(Treatment, "REF")) %>% # ignore REF files
  unnest(processed_spectra) %>% # use corrected spectra
  select(-(ChB:raw_reflectance)) %>%  # remove unnecessary Wavelength specific rows 
  filter(Wavelength > 400 & Wavelength < 1000) %>% 
  filter(corrected_reflectance > 1) %>% 
  nest(Spectra = c(Wavelength, corrected_reflectance))

## File List 
absurd_files <- absurd_data$spu_filename

```

## Apply references

##### Check Application
```{r}
## Check all files have a corrected reflectance
corrected_spectra_files <- df_corrected %>% unnest(processed_spectra) %>% filter(!is.na(corrected_reflectance)) %>% select(spu_filename, Date, Site) %>% distinct()

## missing corrections
anti_join(df_corrected, corrected_spectra_files) %>% 
  group_by(Date, Site, Block, Treatment, Integration) %>% 
  summarize(Num_Files = n()) %>% kable()

## Check that no important treatments are left out 
anti_join(df, df_corrected, by = "spu_filename") %>% pull(Treatment) %>% unique()
```



### Calculate INDICES for new spu files
```{r}
new_files <- spu_dataframe_new$spu_filename

df_processed_new <- df_processed %>% filter(spu_filename %in% new_files)

indices_new <- unispec_file_key_fix %>%
  filter(!is.na(Treatment)) %>%
  filter(!str_detect(Treatment, "REF|DARK")) %>%
  inner_join(df_processed_new) %>%
  ## Format for calculuating indices
  unnest(Spectra) %>%
  select(-ChB, -ChA, -raw_reflectance, -CorrectionFactor) %>%
  rename(Reflectance = corrected_reflectance) %>%
  nest(Wavelength, Reflectance,.key = Spectra) %>%
  ## Calculate NDVI
  mutate(Indices = map(Spectra, function(x) calculate_indices(x, band_defns = band_defns, instrument = "MODIS", indices = c("NDVI", "EVI", "EVI2"))))

## SAVE UPDATED INDICES
indices_all <- bind_rows(df_indices, indices_new)
index_filename <- paste0(data_path, dir_year, "_index_data.rds")
write_rds(indices_all, index_filename)

```


### Label Unispec Problems
```{r unispec_problem_key}

problem_files <- c(maxed_files, dim_files, absurd_files, missing_spu_files)

df_ref_filenames <- df_corrected %>% unnest(processed_spectra) %>%
  select(spu_filename, ref_filenames) %>% distinct()

unispec_problem_key_pre <- unispec_file_key_fix%>%
  mutate(Replicate = factor(Replicate)) %>%
  # mislabeled
  mutate(mislabeled = !is.na(key_fix)) %>% # mislabeled
  # mismeasurement
  left_join(maxed_data) %>% # maxed_number, maxed_wavelenghs
  mutate(maxed = spu_filename %in% maxed_files) %>%
  mutate(dim = spu_filename %in% dim_files) %>% # dim
  mutate(absurd_reflectance = spu_filename %in% c(zero_files, absurd_files)) %>% # absurd_reflectance
  # missing
  mutate(missing_spu = spu_filename %in% missing_spu_files) %>%  # missing_spu
  left_join(df_ref_filenames)

# miscorrection
ref_problem_key <- unispec_problem_key_pre %>%
  filter(spu_filename %in% ref_files) %>% # select ref files
  #mutate(mislabeled = if_else(spu_filename %in% ref_mistakes, TRUE, mislabeled)) %>%
  gather("problem", "status", mislabeled, maxed, dim, absurd_reflectance, missing_spu) %>% # row for each type of problem
  filter(!is.na(status)) %>% # remove non-problems
  filter(status != FALSE) %>% # remove non-problems
  group_by(spu_filename) %>%
  select(spu_filename, problem, status) %>%
  summarize(problems = str_c(unique(problem), collapse = ", ")) %>% # group: one row per spu_filename
  rename(ref_filenames = spu_filename, ref_problem = problems) # rename variables to join w/unispec_problem_key

## Unispec Problem Key
unispec_problem_key <- unispec_problem_key_pre %>%
  separate_rows(ref_filenames, sep = ", ") %>% # split ref_filenames string into multiple rows per ref_filename
  left_join(ref_problem_key) %>% # adds ref_problem column
  group_by(spu_filename) %>%
  summarize(ref_problems = str_c(str_replace_na(ref_problem), collapse = "; ")) %>% # collapse to one row per spu_filename
  right_join(unispec_problem_key_pre) %>%  # add the rest of the key info back in
  mutate_at(.vars = c("ref_problems", "dim", "absurd_reflectance", "mislabeled", "missing_spu"), .funs = factor) %>%
  mutate(file_problem = spu_filename %in% problem_files) %>%
  select(spu_filename, Type, Date, Site, Block, Treatment, Replicate, Weather, Notes, missing_spu, file_problem, mislabeled, key_fix, maxed, maxed_number, maxed_wavelengths, dim, absurd_reflectance, ref_problems, ref_filenames) # order


### Summary table
unispec_problem_key %>%
  mutate_at(list(factor), .vars = vars(c("Site", "Block", "Treatment", "key_fix", "ref_problems"))) %>%
  mutate(file_problem = spu_filename %in% problem_files) %>%
  summary()
```

