---
title: "Process Unispec Files collected on One Day"
author: "Ruby An"
date: "2021-06-24"
output: html_document
editor_options: 
  
  chunk_output_type: console
params:
  
  session_date: 2020-07-16
  data_path: /home/ruby/UnispecData/2020/raw_data/2020-07-16 # select folder where spu files reside
---

## Introduction
This file is fo  processing unispec data in a user friendly way. 

## Required Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F)

## Required Packages
library("tidyverse")
library("knitr")
library(rChoiceDialogs)
library(lubridate)

## Functions file  ## Decide to include functions in processing session for now. 
```

## R Functions
```{r functions, include = F}
read_spu_file_metadata <- function(filename, info = "short") {
  # DESCRIPTION: Reads first 9 text lines in .spu files 
  # INPUT: .spu file -- collected using PPSystems UnispecDC
  #         info = "short" or "long" 
  # OUTPUT: dataframe with 15 columns on instrument status, scan settings, max/min value

    # Extract info from the file itself, reading metadata from first 9 lines. Create a dataframe
    text <- read_lines(filename, n_max=9)
    
    # Line 1: Extract the file name in the spu file as a check. Some file names have spaces 
    spu_filename <- str_replace(text[1],".*[\\\\]([A-z0-9.]\\s*)","\\1") %>% # extract filename
      str_replace("\"","") # removes trailing quote at end of line 
    FileNum <- str_extract(spu_filename, "\\d{5}") %>% as.numeric() # from 5 digits in filename
    
    # Line 2: 
    Remarks <- str_split(text[2], pattern = " ")[[1]] # split by "space"
    Type <- str_split(Remarks[7], pattern = "=")[[1]][2] # extract relevant part
    ScanType <- ifelse(grepl("DARKscan",Type, fixed=T), "DARKscan", # format
                       ifelse(grepl("Datascan,DC",Type, fixed=T), "Throwawayscan", 
                              Type))
    DarkscanID <- str_extract(text[2], "Dark=.+.spu")
    Remarks <- text[2]
    
    # Line 3: 
    DateTime <-  lubridate::mdy_hms(text[3], tz="America/Anchorage")
    
    # Line 4: Limits -- range of spectra measured
    Limits <- str_extract(text[4], "\\d+.\\d.+\\d") 
    
    # Line 5: 
    Temperature <- as.numeric(strsplit(strsplit(text[5], split = " ")[[1]][4], split="=")[[1]][2])
    Battery <- str_extract(text[5], "BattV=\\d+.\\d+")
    Aux <- str_extract(text[5], "A\\d=.+\\d")
    
    # Line 6-9: 
    Minimum <- str_extract(text[6], "\\d+.+\\d") # Wavelength, ChB min
    Minimum_wavelength <- str_split(Minimum, boundary("word"))[[1]][1] # Wavelength
    Minimum_value <- str_split(Minimum, boundary("word"))[[1]][2] # ChB AD
    
    Maximum <- str_extract(text[7], "\\d+.+\\d") # Wavelength, ChB max
    Maximum_wavelength <- str_split(Maximum, boundary("word"))[[1]][1]
    Maximum_value <- str_split(Maximum, boundary("word"))[[1]][2]

    Integration <- as.numeric(strsplit(text[8], split = " ")[[1]][3])
    NumberScans <- str_extract(text[9], "\\d+")
    
    # Truncated Filename - use as SCANID to join to other dataframes
    spu_filename <- unlist(str_split(filename, pattern = "/")) %>% last()
    
    # Metadata 
    metadata <- tibble(spu_filename, DateTime, FileNum, ScanType, Integration, NumberScans,Minimum_wavelength,Minimum_value, Maximum_wavelength, Maximum_value, Limits, Temperature, Battery, Aux, DarkscanID, Remarks)
    
    if(info == "short") {
      metadata <- metadata %>% select(spu_filename, DateTime, FileNum, Integration)
    }

  # Print filenames while reading 
  #print(spu_filename) # use for error checking
  
  return(metadata)
}


read_spu_file_spectra <- function(filename) {
  # DESCRIPTION: For a generic .spu file regardless of name, extract spectral data
  # INPUT: Unispec-DC .spu file
  # OUTPUT: dataframe of spectral data with 3 columns Wavelength, ChB, ChA
  
  # Read spectral intensity data into dataframe
  data <- read.table(file = filename, skip = 9, col.names = c("Wavelength", "ChB", "ChA"))
  
  print(filename)
  
  return(data)
}

# Color band definitios for calculate_indices function
band_defns <- tribble(
  ~definition, ~color, ~min, ~max,
  "ITEX", "red", 560, 600,
  "ITEX", "nir", 725, 1000,
  "MODIS", "red", 620, 670, 
  "MODIS", "nir", 841, 876,
  "MODIS", "blue", 459,479,
  "SKYE", "red", 620, 680,
  "SKYE", "nir", 830, 880,
  "SKYE", "blue", 455, 480,
  "ToolikGIS_Drone_2018", "red", 640, 680,
  "ToolikGIS_Drone_2018", "nir", 820, 890,
  "ToolikGIS_MicaSense_2019", "blue", 455, 495,
  "ToolikGIS_MicaSense_2019", "green", 540, 580,
  "ToolikGIS_MicaSense_2019", "red", 658, 678,
  "ToolikGIS_MicaSense_2019", "red_edge", 707, 727,
  "ToolikGIS_MicaSense_2019", "near_ir", 800, 880,
  "ToolikEDC", "red", 560, 680,
  "ToolikEDC", "nir", 725, 1000
)

calculate_indices <- function(spectra, band_defns, instrument = "MODIS", indices = "NDVI") {
  # Calculates NDVI, EVI, and EVI2 from dataframe including Wavelength : Spectra 
  ## inputs: spectra - Wavelength, Reflectance columns
  ##         band_defns : wavelengths definining colors 
  ##         instrument : e.g. MODIS, SKYE, ITEX
  ##         indicies   : the index to return 
  ## output: Index - name of vegetation index
  ##         BandDefinition - name of "instrument" or spectral band definition used
  ##         Value - value of index, with the band definition used. 
  
  bands <- band_defns %>% 
    filter(definition == instrument) 
  
  blue <- bands %>% filter(color=="blue") %>% select(min, max) %>% as.numeric()
  nir <- bands %>% filter(color=="nir") %>% select(min, max) %>% as.numeric()
  red <- bands %>% filter(color=="red") %>% select(min, max) %>% as.numeric()
  
  spectra_bands <- spectra %>% 
    mutate(color = ifelse(Wavelength >= blue[1] & Wavelength <= blue[2], "blue",
                          ifelse(Wavelength >= red[1] & Wavelength <= red[2], "red",
                                 ifelse(Wavelength >= nir[1] & Wavelength <= nir[2], "nir",
                                        "other")))) %>% 
    group_by(color) %>% 
    summarize(Reflectance = mean(Reflectance))
  
  index_data <- spectra_bands %>%
    spread(color, Reflectance) %>% 
    mutate(NDVI = (nir-red)/(nir+red),
           EVI = 2.5*((nir-red)/(nir+6*red-7.5*blue + 1)),
           EVI2 = 2.5*((nir-red)/(nir+2.4*red + 1))) %>% 
    select_at(indices) %>% 
    gather(Index, Value, everything()) %>% 
    
    # Add Spectral Band Definition convention
    mutate(BandDefinition = instrument) %>% 
    select(Index, BandDefinition, Value)
  
  return(index_data) 
}

```

## Digitize Data Locally / Renaming raw files
Rename files to include parent directory (Date) in filename: "YEAR-MONTH-DAY_SITE_FileNum.spu". Copy the following code and use in the command line (haven't figured out how to run in RStudio yet... some kind of error that "file doesn't exist"). Adjust so that year (date) structure is correct. 
```{bash}

cd /UnispecData/2020  # go to directry with session "date folders"
#Set $SESSION = session date somehow 

#mkdir raw_data

# RUN THIS TO RENAME FILES AND MOVE THEM INTO .spu folders 
#!/bin/bash
for date in 2020-07-19; do
  pushd "$date"
  for file in *; do
    cp "$file" ../"${date}_${file}"
  done
  popd
done

  mv *.spu raw_spectra_files/ #make sure the "raw_data" folder exists first! :) 

```

## Data directories, date and key files

  * Session_date in yyyy-mm-dd form.
  
  * data_path is year folder where the session files are stored. The structure is year/spu/yyyy-mm-dd  where yyyy-mm-dd is sample date. Should be on OneDrive.
  
  * Field Key file with header "Date,Site,Block,Treatment,Replicate,Location,FileNum,Notes,Weather".  File name convention is year_unispec_field_key.csv
  
```{r, include = F}
## Session Date

session_year <- year(params$session_date)

## use the params$data_path as starting folder, this should be to the raw_files 
data_path <- params$data_path

```

### Read .SPU files 
  
```{r, include = F}
## Make a character vector of all .spu raw files in current session date folder
# Check if directory exist and use it as the working directory

spu_files <- list.files(path = data_path, pattern = ".spu$", full.names = T, recursive=T)

# Read metadata text lines (9) from the spu files
spu_filedata <- map_dfr(spu_files, read_spu_file_metadata)

# Add Metadata columns:
#   spu_filename_full: variable with the full path filename 
#   Site: column from spu_filename
#   FileNum: column from spu_filename 
#   ScanType: column identifying scan type (Dark, Throwaway, Datascan)
spu_metadata <- spu_filedata %>% 
  mutate(spu_filename_full = spu_files) %>% #
  mutate(Site = toupper(str_extract(spu_filename, "[A-Za-z]{3}[0-9]{1,2}(?=_)"))) %>% # get string that is 3 letters, 2 numbers before _ 
  mutate(FileNum = as.integer(str_extract(spu_filename, "\\d{5}"))) 

# Read spectra from .spu files > add to metadata
spu_dataframe <- spu_metadata %>% 
  mutate(Spectra=map(spu_filename_full, function(x) read_spu_file_spectra(x))) %>% 
  mutate(Date = date(DateTime))
  # mutate(Time = DateTime-floor_date(DateTime, unit = "day")) 

```

### Read Unispec File Key
```{r}

key_files <- list.files(path = data_path, pattern = "*unispec_key.csv$", full.names = T, recursive=T)

field_keys <- key_files %>%  purrr::map(function(file_name) read_csv(file_name)) %>% 
  reduce(rbind) %>% 
  mutate(Date = date(params$session_date)) # Add date


# Join by SITE, DATE, FILENUM
df <- left_join(spu_dataframe, field_keys) %>% arrange(DateTime) %>% 
  mutate_at(.vars = vars(Site, Block, Treatment), .funs = factor)
```

## REF Correction

### Confirm REF choice

White references correct for instrument & cable irregularities. Multiplying by the correction factor (ChA/ChB) smooths out the spectra. There are typically 5 reference measurements per *Date* / *Site*. If multiple file numbers are listed, the correction factors are averaged. 

Based on the original field notebook key and the following quality checks, choose reference files by entering the appropriate file numbers in **`r field_key`** for the rows where the column *Treatment* = **REF**. 

Make sure to rerun the **Load Keys**. The following plots your chosen references.

### Extract REF Data
```{r ref_table, echo=F}
options(knitr.kable.NA = '')

## Build REF key 
ref_keys <- field_keys %>% 
  filter(Treatment == "REF") %>% # extract reference data 
  ## The following separates the Site column into "Site" and "Site_REF"
  ### Site = the site to which the reference measurements should be applied 
  ### Site_REF = the site where the reference measurements were actually collected
  separate(Site, into=c("Site", "Site_REF"), sep = "_", fill="right") %>% 
  mutate(Site_REF = coalesce(Site_REF, Site)) # if the references were collected at 'Site', the created column Site_REF will be NA. Use coalesce() to fill these NA's with the value of "Site".  

### spu data for references
spu_for_ref <- spu_dataframe %>% # in "spu_dataframe", the "Site" column is the location where the data was collected 
  rename(Site_REF = Site) # we thus rename Site to Site_REF to match the column 'ref_keys'

## Join spu data to ref_keys by Site_REF, Date, FileNum
ref_data <- ref_keys  %>% 
  left_join(spu_for_ref)  ## "Site_REF" is the location where the file (from which the reference correction factor is calculated) actually was collected
  ## "Site", inherited from ref_keys, is now the location where the correction factor should be applied


## Table per Site of Reference Files 
ref_filenames_table <- ref_data %>% group_by(Date, Site, Site_REF) %>% 
  summarize(Files = n_distinct(spu_filename), ref_filenames = str_c(spu_filename,collapse = ", ")) 

## Table per Site for all files 
df_ref_table <- df %>% 
  separate(Site, into=c("Site", "Site_REF"), sep = "_", fill="right") %>%
  filter(!is.na(Treatment)) %>% 
  group_by(Date, Site) %>% 
  summarize(Treatments = str_c(unique(Treatment), collapse = ","), Files = n_distinct(spu_filename))

## Output
ref_filenames_table
```


### VIZ CHECK: Plot References
```{r ref_choice_list}

ref_data %>% filter(Spectra %>% map(is.null) %>% map_lgl(any)) %>% select(Date, Site, Block) %>% unique()

## Build Plot all reference data files
ref_data_all <- ref_data %>% unnest(Spectra) %>%
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  mutate(CorrectionFactor = ChA/ChB)

(ref_plot_all <- ggplot(ref_data_all, aes(x = Wavelength, y = CorrectionFactor, group=spu_filename)) + 
  theme(legend.position="left") + 
  geom_line(aes(color=factor(Integration))))

## File Lists
ref_files <- ref_data$spu_filename %>% unique()

```

### Apply References
Rerun the **Join Data & keys** sections above to update the `df_clean` dataframe.Apply your chosen references to actual spectral data to create the tidy dataframe `df_tidy` containing corrected sepectral reflectance values.

```{r apply_refs, echo=FALSE}
## Average 5 chosen ref measurements per DATE/SITE/BLOCK 
ref_summary <- ref_data %>% 
  ## The following steps expand the "Block" column to create one REF set per Block per Site. This structure is necessary for situtations where different refs are used for different blocks at the same site. 
  separate(Block, into = c("BX1", "BX2", "BX3", "BX4"), sep = ",") %>% #1: expand string entry in "Block" into separate columns -- NOTE: this step throws a "Warning: Expected 4 pieces." for sites w/less than 4 blocks
  gather(Block, BlockString, BX1:BX4) %>% #2: re-condense into one column, generates correct number of rows per site AND per block
  mutate(Block = str_squish(BlockString), BlockString=NULL) %>% #3: replace placeholder column names w/"B1-B4". Also removes whitespace from BlockString contents introduced by "separate" function
  filter(!is.na(Block)) %>% #4: remove empty rows for sites w/out B3 or B4
  ### Unnest Spectra & calculate
  unnest(Spectra) %>% 
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  mutate(CorrectionFactor = ChA/ChB) %>% 

  ### The following code group repeated REF measurements, and takes the mean 
  group_by(Date,Site,Block,Integration, Wavelength) %>% 
  # group_by(Date,Site,Block,Wavelength, Integration_ms) %>% # to separate integration times
  summarize(ChA_REF = mean(ChA), ChB_REF = mean(ChB), CorrectionFactor = mean(ChA/ChB), Notes_REF = str_c(Notes, collapse = "; "), ref_filenames = str_c(spu_filename,collapse = ", "))

## Join DATA with REFS
### Check no missing Spectra
# spu_dataframe %>% filter(Spectra %>% map(is.null) %>% map_lgl(any))

spu_for_plots <- df %>% filter(!str_detect(Treatment, "REF|DARK")) %>% filter(!is.na(DateTime)) %>% 
  unnest(Spectra) %>% 
  filter(Wavelength > 400, Wavelength < 1000)


df_ref <- left_join(spu_for_plots, ref_summary) %>%
  mutate(raw_reflectance = ChB/ChA) %>% # the raw reflectance
  mutate(corrected_reflectance = raw_reflectance*CorrectionFactor) 


## Corrected Reflectances 
df_corrected <- df_ref %>% 
  nest(processed_spectra = c(Wavelength, ChB, ChA, ChB_REF, ChA_REF, CorrectionFactor, raw_reflectance, corrected_reflectance))
```


##### Plot Check
```{r}

(df_corrected %>% 
  slice(1:10) %>% 
  unnest(processed_spectra) %>% 
  gather(key = Status, value = Reflectance, raw_reflectance, corrected_reflectance) %>% 
  
  # VIZ
  ggplot(mapping = aes(x = Wavelength, y = Reflectance )) + 
  geom_line(aes(color = spu_filename, linetype = Status)) + 
    facet_wrap(vars(Date, Site, Block, Treatment, FileNum)))

```


## ? INTERPOLATION


## SAVE DATA
```{r}
processed_spu_data <- df_corrected %>%
  select(spu_filename, DateTime, ref_filenames, processed_spectra) %>% unnest(processed_spectra) %>%
  select(spu_filename, DateTime, ref_filenames, Wavelength, ChB, ChA, raw_reflectance, CorrectionFactor, corrected_reflectance) %>%
  nest(Spectra = c(Wavelength, ChB, ChA, raw_reflectance, CorrectionFactor, corrected_reflectance))

ref_spu_data <- ref_data %>% unnest(Spectra) %>%
  select(spu_filename, DateTime, Wavelength, ChB, ChA) %>%
  mutate(Reflectance = ChB/ChA) %>% 
  mutate(CorrectionFactor = 1/Reflectance, corrected_reflectance = NA) %>%
  filter(Wavelength > 400, Wavelength < 1000) %>%
  rename(raw_reflectance = Reflectance) %>%
  nest(Spectra = c(Wavelength, ChB, ChA, raw_reflectance, CorrectionFactor, corrected_reflectance))

processed_spectra <- bind_rows(processed_spu_data, ref_spu_data)

## Save
processed_spu_filename <- paste0(data_path, "/", session_year, "_processed_spu_data.rds")

write_rds(df_processed, processed_spu_filename)
```


## Calculate INDICES

Currently, this only works for NDVI, EVI, and EVI2 as I haven't worked out spectral interpolation yet and the other indices need reflectance at a specific value (not a range).

### Calculate Indices
```{r}

df_processed %>%  filter( Spectra %>% map(is.null) %>% map_lgl(any)) # remove files w/out spu_spectra

df_indices <- df %>% select(-Spectra) %>% #dataframe with metadata
  filter(!is.na(Treatment)) %>%
  filter(!str_detect(Treatment, "REF|DARK")) %>%
  inner_join(df_processed) %>% # corrected reflectances with 
  ## Format for calculuating indices
  unnest(Spectra) %>%
  select(-ChB, -ChA, -raw_reflectance, -CorrectionFactor) %>%
  rename(Reflectance = corrected_reflectance) %>%
  nest(Spectra = c(Wavelength, Reflectance)) %>%
  ## Calculate NDVI
  mutate(Indices = map(Spectra, function(x) calculate_indices(x, band_defns = band_defns, instrument = "MODIS", indices = c("NDVI", "EVI", "EVI2"))))

index_filename <- paste0(data_path, "/", params$session_date, "_index_data.rds")
write_rds(df_indices, index_filename)


## convert to SHINY ready format


```


### Visualize with other years

### Check across years

## CHECKLIST/SUMMARY REPORT



