---
title: "Process Unispec Files"
author: "Ruby An"
date: "6/17/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Required Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Required Packages
library("tidyverse")
library("knitr")
source("UnispecProtocol/unispec_protocol_functions.R") # file loads required functions

## Session Date
session_date <- "2019-08-05"

## for Ruby's chromebook 
data_path <- paste0("UnispecRecord/UnispecData/2019_unispec_data/", session_date, "/")

## use data stored outside of git directory 
data_path <- paste0("/home/ruby/UnispecData/2019_unprocessed_unispec_data/", session_date, "/")



```


## Digitize Data Locally 

### Read in spu_files 
  
```{r}
## Make a character vector of all .spu raw files in folder
spu_files <- list.files(path = data_path, pattern = ".spu$", full.names = T, recursive=T)

# Read metadata text lines (9) from the spu files
spu_filedata <- map_dfr(spu_files, read_spu_file_metadata)

# Add Metadata columns:
#   spu_filename_full: variable with the full path filename 
#   Site: column from spu_filename
#   FileNum: column from spu_filename 
#   Type: column identifying scan type (Dark, Throwaway, Datascan)
spu_metadata <- spu_filedata %>% 
  mutate(spu_filename_full = spu_files) %>% 
  mutate(Site = str_extract(spu_filename, "([A-Z]{3,8}[0-9]*)(?=_)")) %>% # get string after last / & before _ 
  mutate(FileNum = as.integer(str_extract(spu_filename, "\\d{5}"))) %>% # extract 5 digits
  mutate(Type = ifelse(grepl("DARKscan",Remarks, fixed=T), "Darkscan",
                       ifelse(grepl("Datascan,DC",Remarks, fixed=T), "Throwawayscan", NA))) %>% 
  distinct(DateTime, spu_filename, .keep_all = T)# %>% 
  # mutate(Date = as.character(Date))

# Read spectra from .spu files > add to metadata
spu_dataframe <- spu_metadata %>% 
  mutate(Spectra=map(spu_filename_full, function(x) read_spu_file_spectra(x)))


## Data summary
spu_dataframe %>%   group_by(Date, Site) %>% 
  summarize(Files = n_distinct(spu_filename)) %>% 
  kable()

```


### Read Field Key 
```{r}

field_key_files <- list.files(path = data_path, pattern = "key.csv$", full.names = T, recursive=T)

field_keys <- map_dfr(field_key_files, read_csv)


## quick cheat: use full field key for 2019
field_key <- "/home/ruby/UnispecData/2019_unprocessed_unispec_data/2019_unispec_field_key.csv"

field_keys <- read_csv(field_key, skip_empty_rows = T) %>% ## remove NA rows 
  filter(Date == session_date)

# Confirm spu_dataframe has same Site abbreviation as key

# > Standardize Site Names -----------------------------------------------------
# Check for different spelling of site names

unique(spu_dataframe$Site) %>% sort()
# spu_dataframe$Site <- recode (spu_dataframe$Site, !!!Site_Names, .default = spu_dataframe$Site)
unique(unispec_field_key$Site) %>% sort()

df <- left_join(spu_dataframe, field_keys) %>% arrange(DateTime) %>% 
  mutate_at(.vars = vars(Site, Block, Treatment), .funs = factor)


```


##### Missing files 
List missing raw .spu files and those that are unprocessed (raw .spu file does not appear in summary .xlsx). Missing raw files
```{r missing}
### MISSING: Find all processed data that is missing corresponding raw spu files 
missing_spu_data <- anti_join(field_keys, spu_dataframe) %>% 
  filter(!str_detect(Site, "_"))

missing_spu_data %>% group_by(Date, Site) %>% 
  summarize(Treatments = str_c(unique(Treatment), collapse = ","), Files = n()) %>% 
  kable()


(unprocessed_spu_files <- anti_join(spu_dataframe, field_keys))
```


## Process Data 

### QAQC 

#### Instrument Check
```{r check_refs}

first_file <-0
last_file <- 15
check_site <- "DHT"

df %>% 
  #filter(Date == check_date) %>% 
  filter(Site %in% check_site) %>% 
  filter(FileNum >= first_file) %>%
  filter(FileNum <= last_file) %>% 
  
  # arrange for 
  unnest(Spectra) %>% 
  filter(FileNum >= first_file) %>%
  filter(FileNum <= last_file) %>%
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  mutate(Reflectance = ChB/ChA) %>% 
  gather(key = Channel, value = Intensity, ChB, ChA) %>% 
  gather(key = ref_part, value = Reflectance_Intensity, Intensity, Reflectance) %>% 
  
  # Viz
  ggplot(mapping = aes(x = Wavelength, y = Reflectance_Intensity)) +
  geom_line(aes(color=Integration_ms, linetype=Channel)) +
  facet_grid(ref_part ~ Date + Site + FileNum + Integration_ms, scales="free") 



```


###### Max'd Out Spectra (> 65000 AD)
List files that maxed out, in the wavelengths used to calculate MODIS NDVI.
```{r maxed, dependson="df"}

maxed_data <- df %>% inner_join(spu_dataframe %>% select(spu_filename)) %>% # restrict to those w/.spu files
  unnest(Spectra) %>% 
  filter(ChA > 65000 | ChB > 65000) %>% 
  group_by(spu_filename) %>% 
  summarize(maxed_number = n(), maxed_wavelengths = str_c(Wavelength, collapse = ", "))  

# Print MAXED files 
if (nrow(maxed_data) > 0) {
  maxed_data %>% inner_join(df) %>% group_by(Site, Date) %>% 
  summarize(Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), 
            Num_Files = n(), 
            Num_Maxed = round(mean(maxed_number), digits = 1)) %>%
  kable()
  
  maxed_files <- maxed_data$spu_filename
  
  
  ## Select those Max'd in NDVI Region
maxed_limit <- 5 # keep files where spectra is max'd only at a narrow peak (e.g. for maxed_limit = 5, maxed region < 5*3.3nm = 16.5 nm)

maxed_data_bad <- df %>% inner_join(spu_dataframe %>% select(spu_filename)) %>% unnest(Spectra) %>% 
  filter(ChA > 65000 | ChB > 65000) %>% 
  ## Extra conditions for NDVI region
  filter(Wavelength > 620) %>% # MODIS red lower bound
  filter(Wavelength < 876) %>% # MODIS nir upper bound
  ## Summary
  group_by(spu_filename) %>% 
  summarize(maxed_number = n(), maxed_wavelengths = str_c(Wavelength, collapse = ", ")) %>% 
  filter(maxed_number > maxed_limit)

## Summary of NDVI Max'd files 
maxed_data_bad %>% inner_join(df) %>% group_by(Site, Date) %>% 
  summarize(Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), Num_Files = n(), 
            Num_Maxed = round(mean(maxed_number), digits = 1)) %>% 
  print(n=50)

maxed_files_bad <- maxed_data_bad$spu_filename
  
  
} else { print("no maxed files")}



```


##### Dim'd Out Spectra
Primarily Darkscans should show up. 
```{r dim, dependson=df}
dim_data <- df %>% inner_join(spu_dataframe %>% select(spu_filename)) %>% unnest(Spectra) %>% 
  group_by(spu_filename) %>% 
  summarize(ChA_max = max(ChA)) %>% 
  filter(ChA_max < 20000) %>% 
  left_join(df) 

# Summary of important info
dim_data %>% group_by(Date, Site) %>% 
  filter(!is.na(Treatment)) %>% 
  filter(!str_detect(Treatment, "DARK|Throwawayscan")) %>% 
  summarize(Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), 
            Types = str_c(str_replace_na(unique(Type)), collapse = ","), 
            Files = n(), ChA_Max = max(ChA_max))  %>% 
  kable()

# File List
dim_files <- dim_data$spu_filename
```


###### Zero'd Spectra
```{r zerod}
## zero'd data 
zero_data <- df %>% inner_join(spu_dataframe %>% select(spu_filename)) %>% unnest(Spectra) %>% 
  filter(ChA == 0) %>%  ## This is for all wavelengths, not just 400-1000nm
  group_by(spu_filename) %>% 
  summarize(Zeros = n()) %>% 
  left_join(df)
  
## Summary 
zero_data %>% group_by(Date, Site) %>% 
  summarize(Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), 
            Files = n(),
            Max_Zeros = max(Zeros)) %>% 
  kable()


## File List : Restricted Wavelengths
zero_files <- zero_data %>% unnest(Spectra) %>% 
  filter(ChA == 0 ) %>% 
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  pull(spu_filename) %>% unique()

zero_data_narrowed <- df %>% filter(spu_filename %in% zero_files) 

## Plot Exploration
zero_data_narrowed %>% 
  slice(1:5) %>% plot_channels()

## Summary
zero_data_narrowed %>% group_by(Date, Site) %>% 
  summarize(Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), Files = n())  %>% 
  kable()
  
```

##### Reflectance >1 
```{r absurd}

## Raw Spectra
absurd_raw_data <- df %>% 
  filter(! Spectra %>% map(is.null) %>% map_lgl(any)) %>% # remove files w/out spu_spectra
  filter(!str_detect(Treatment, "REF|DARK")) %>% # ignore REF files
  unnest(Spectra) %>% # use raw spectra
  filter(Wavelength > 400 & Wavelength < 1000) %>% 
  filter(Reflectance > 1) %>% 
  nest(Wavelength, ChA, ChB, Reflectance, .key = Spectra)

absurd_raw_files <- absurd_raw_data$spu_filename

absurd_raw_data %>% group_by(Date, Site) %>% 
  summarize(Treatments = str_c(unique(Treatment), collapse = ","), Files = n())

## Corrected Spectra
absurd_data <- df_corrected %>% 
  filter(!str_detect(Treatment, "REF")) %>% # ignore REF files
  # filter(!spu_filename %in% unprocessed_spu_files) %>% # leave out missing multispec data
  unnest(processed_spectra) %>% # use corrected spectra
  select(-(ChB:raw_reflectance)) %>%  # remove unnecessary Wavelength specific rows 
  filter(Wavelength > 400 & Wavelength < 1000) %>% 
  filter(corrected_reflectance > 1) %>% 
  nest(Wavelength, corrected_reflectance, .key = corrected_spectra)

## Summary
absurd_data %>% group_by(Date, Site) %>% 
  summarize(Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), Files = n())  %>% 
  kable()

## File List 
absurd_files <- absurd_data$spu_filename

## Plot Check 
absurd_data %>% 
  inner_join(spu_dataframe) %>% 
  slice(1:10) %>% 
  plot_channels()

```

#### Reality Check: Times between files, Mislabeling, etc.
```{r time_check}

## SPECIFY SITE/DATE/ETC to ZOOM IN ON
check_site <- "LMAT"
check_date <- "2019-06-02" # necessary to unlist dates vector

## Files 
first_file <- 0
last_file <- 153

# Select columns to Check
timedata <- df %>% 
  #filter(Date == check_date) %>% 
  #filter(Site %in% check_site) %>% 
  filter(FileNum >= first_file) %>%
  filter(FileNum <= last_file) %>% 
  select(Site, Date, DateTime, FileNum, Integration_ms, Type) %>% 
  distinct(DateTime, .keep_all = T)

timedata$diff <- timedata$DateTime - lag(timedata$DateTime)

meta_timedata <- left_join(timedata, field_keys)

time_check <- meta_timedata %>% select(Site, Date, DateTime, Block, Treatment, Replicate, Location, FileNum, diff, Integration_ms, Type) %>% ungroup()

# Examine dataframe
time_check %>% select(Date, Site, Block, Treatment, Location, diff, Integration_ms) %>% print(n=200)
```

##### Mis-Labeling 
Check for human error in recording correspondance between metadata and file numbers. 

###### Discrepancy Checks
```{r df_check, dependson="df", eval= T, echo=F}
##### ------------------------- Variable checks 

# Date inconsistencies
spu_dataframe %>% mutate(Date_unispec = lubridate::date(DateTime)) %>% 
  filter(Date != Date_unispec)  %>% 
  select(Site, Date, Date_unispec) %>% distinct() %>%
  print(n=50)


# Missing spu files 
df %>% filter(!str_detect(Site, "_")) %>% 
  filter(is.na(spu_filename))  %>% 
  select(spu_filename, Site, FileNum, Date, Block, Treatment, Replicate, Notes, key_fix) %>% 
  group_by(Date, Site) %>% 
  summarize(Files = n()) %>% 
  kable()
  

# Missing DateTime (equivalent to spu file)
df %>% filter(is.na(DateTime)) %>% 
  filter(!str_detect(Site, "_")) %>% 
  select_if(function(x) typeof(x) != "list") %>% summary()


# Block NA's (if in key)
df %>% filter(!is.na(Treatment)) %>% filter(is.na(Block)) %>% 
  filter(!str_detect(Treatment, "DARK|REF|VEG"))

df %>% filter(str_detect(Treatment, "REF"))  %>% 
  filter(is.na(Block)) %>% print(n=100)

# Replicate NA
df %>% filter(!is.na(Treatment)) %>% filter(is.na(Replicate)) %>% 
  filter(!str_detect(Treatment, "DARK|REF|VEG"))

## General NA test
df_na <- df %>% filter(is.na(spu_filename) |
              is.na(Site) |
              is.na(Block) & !str_detect(Treatment, "REF|DARK|VEG") | # Block NA's should always be REFS or EXTRA
              is.na(Replicate) & !str_detect(Treatment, "REF"), # Check for replicate NA's that aren't REF
              Treatment != "EXTRA|VEG|REF") # don't care about EXTRA

df_na %>% filter(Treatment !="REF") %>% select_if(function(x) typeof(x) != "list") %>% summary()

#####
## Check for large filenumbers : Inconsistent File Number reading due to number at end of site names
df_lgfn <- df %>%
  filter(FileNum > 600) %>%  # Sites w/numbers at the end mis-read in as part of file number. All > 100000: Unispec-DC measurements only go up to 5 digits.
  select(Date, Site, Block, Treatment, FileNum, spu_filename)
df_lgfn %>% print(n=50)


#####
## Check Site Names
df_sites <- unique(df$Site)
df_sites

## Check Treatment names
df_treatments <- unique(df$Treatment)
df_treatments

### -------------------------------------- SUMMARY CHECKS

## Dataframe 
df_summary <- df %>% group_by(Site, Date) %>%  
  filter(!is.na(Treatment)) %>% 
  summarize(Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), Num_Files = n()) 

df_summary %>% print(n=100)

# Discrepancies  
## Site vs. Site_filename differences 
#### should only be Reference files or mislabeled (e.g. MAT extends into LMAT, or NANT continues to MNAT)
df_site_diff <- field_keys %>% left_join(spu_key %>% select(-Date, -Type, -FileNum) %>% rename(Site_filename = Site)) %>%   filter(Site != Site_filename) 

df_site_diff %>% 
  group_by(Date) %>% 
  summarize(Sites = str_c(unique(Site), collapse = ", "), Sites_filename = str_c(unique(Site_filename), collapse = ", "), Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), Num_Files = n())

## Check File Number Pattern:
#### Num Files per block should be multiples of 5, unless REF or NA
df_filenum_count <- field_keys %>% group_by(Site, Date, Block) %>% 
  filter(Treatment != "DARK") %>% 
  filter(Treatment != "REF") %>% 
  summarize(Treatments = str_c(str_replace_na(unique(Treatment)), collapse = ","), Num_Files = n()) %>% 
  #filter(Treatments != "NA|REF|DARK") %>% 
  filter(Num_Files %% 5 != 0)  # files per plot
  #filter(!str_detect(Treatments, "REF|NA")) 

df_filenum_count

#df %>% inner_join(df_filenum_count) %>% print(n=100) 

```

Run the following code for Discrepancy check output. 
```{r, checks, eval = T, echo=F}

# cat(paste0("Dataframe Columns: ", str_c(df_names, collapse = ", ")))
# cat(paste0("Sites: ", str_c(df_sites, collapse = ", ")))
# cat(paste0("Treatments: ", str_c(df_treatments, collapse = ", ")))
# 
# df_na %>% kable(caption = "NA Metadata")
# df_lgfn %>% kable(caption = "Large FileNumbers")

df_names
df_sites
df_treatments

df_lgfn 
df_site_diff
df_filenum_count

df_summary

### Summarize  - look for NA's
unispec_file_key %>% select_if(function(x) typeof(x) != "list") %>% summary()
df %>% select_if(function(x) typeof(x) != "list") %>% summary()

```

###### Duplicates 
Identify duplicates. Fix by editing unispec_key_fix.csv. Only duplicates should be REF files used for other sites. 

```{r duplicates}
### Check raw spu_files have no duplicates
nrow(spu_dataframe) == length(unique(spu_dataframe$spu_filename))

## Check for duplicates in key
unispec_file_key %>% group_by(Date, Site, Block, Treatment, Replicate, FileNum) %>%
  filter(!(is.na(Site) & is.na(Treatment) & is.na(FileNum))) %>% 
  filter(n()>1) %>%
  arrange(Date, Site, FileNum)

## Check for duplicate in dataframe
duplicates <- df %>% 
  group_by(spu_filename) %>%
  filter(n()>1) %>%
  ungroup() %>%
  arrange(DateTime, spu_filename) %>% 
  select( spu_filename, FileNum, DateTime, Site, Block, Treatment, Replicate, Weather, everything())

 duplicates %>% print(n=200)
# duplicates %>% filter(!str_detect(Treatment, "REF")) %>% arrange(spu_filename) %>% print(n=100)

duplicates %>% 
  #mutate(Date = coalesce(Date, lubridate::date(DateTime))) %>% 
  group_by(Date) %>% 
  summarize(Sites = str_c(unique(Site), collapse = ","), Treatments = str_c(unique(Treatment), collapse = ","), Files  = n()) %>% 
  kable()

duplicates %>% filter(!str_detect(Treatment, "DARK|REF|VEG|TOWER"))

```




### REF Correction

#### Confirm REF choice

White references correct for instrument & cable irregularities. Multiplying by the correction factor (ChA/ChB) smooths out the spectra. There are typically 5 reference measurements per *Date* / *Site*. If multiple file numbers are listed, the correction factors are averaged. 

Based on the original field notebook key and the following quality checks, choose reference files by entering the appropriate file numbers in **`r key_files`** for the rows where the column *Treatment* = **REF**. 

Make sure to rerun the **Load Keys**. The following plots your chosen references.

##### Extract REF Data
```{r ref_table, echo=F}
options(knitr.kable.NA = '')

## Build REF key 
ref_keys <- field_keys %>% 
  filter(Treatment == "REF") %>% # extract reference data 
  ## The following separates the Site column into "Site" and "Site_REF"
  ### Site = the site to which the reference measurements should be applied 
  ### Site_REF = the site where the reference measurements were actually collected
  separate(Site, into=c("Site", "Site_REF"), sep = "_", fill="right") %>% 
  mutate(Site_REF = coalesce(Site_REF, Site)) # if the references were collected at 'Site', the created column Site_REF will be NA. Use coalesce() to fill these NA's with the value of "Site".  

### spu data for references
spu_for_ref <- spu_dataframe %>% # in "spu_dataframe", the "Site" column is the location where the data was collected 
  rename(Site_REF = Site) # we thus rename Site to Site_REF to match the column 'ref_keys'

## Join spu data to ref_keys by Site_REF, Date, FileNum
ref_data <- ref_keys %>% select(Date, Site, Site_REF, FileNum, Block, Treatment, Replicate, Weather, Notes) %>% 
  left_join(spu_for_ref)  ## "Site_REF" is the location where the file (from which the reference correction factor is calculated) actually was collected
  ## "Site", inherited from ref_keys, is now the location where the correction factor should be applied

## Table of Reference Files
ref_table <- ref_data %>% group_by(Date, Site_REF) %>% 
  summarize(Sites = str_c(unique(Site), collapse = ","), Files = n_distinct(spu_filename)) %>% 
  kable()

## Table per Site of Reference Files 
ref_filenames_table <- ref_data %>% group_by(Date, Site, Site_REF) %>% 
  summarize(Files = n_distinct(spu_filename), ref_filenames = str_c(spu_filename,collapse = ", ")) 

## Table per Site for all files 
df_ref_table <- df %>% 
  separate(Site, into=c("Site", "Site_REF"), sep = "_", fill="right") %>%
  filter(!is.na(Treatment)) %>% 
  group_by(Date, Site) %>% 
  summarize(Treatments = str_c(unique(Treatment), collapse = ","), Files = n_distinct(spu_filename))

#### Check that all Date / Sites have references 
df_missing_refs <- df_ref_table %>% 
  filter(!str_detect(Treatments, "REF")) %>%  
  print(n=100)

## Output
ref_table
ref_filenames_table
df_missing_refs


```


##### Plot References
```{r ref_choice_list}

ref_data %>% filter(Spectra %>% map(is.null) %>% map_lgl(any)) %>% select(Date, Site, Block) %>% unique()

## Build Plot all reference data files
ref_data_all <- ref_data %>% unnest(Spectra) %>%
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  mutate(CorrectionFactor = 1/Reflectance)

(ref_plot_all <- ggplot(ref_data_all, aes(x = Wavelength, y = CorrectionFactor, group=spu_filename)) + 
  theme(legend.position="left") + 
  geom_line(aes(color=factor(Integration_ms))))

#+ geom_line(aes(color=factor(Date > "2019-07-02"))) # date


## Build Plot all reference mistakes
ref_data_mistakes <- ref_data_all %>%
  filter(CorrectionFactor > 5) %>% 
  distinct(spu_filename) %>% 
  select(spu_filename) %>% 
  left_join(ref_data_all)

(ref_plot_mistakes <- ggplot(ref_data_mistakes, aes(x = Wavelength, y = CorrectionFactor)) +
  geom_line(aes(color=spu_filename)))

## PLOTS
ref_plot_all + ggtitle("ALL REFERENCES")
# ref_plot_mistakes + ggtitle("REF Mistakes") +
#   scale_y_continuous(limits = c(0, NA))

## File Lists
ref_files <- ref_data$spu_filename %>% unique()
ref_mistakes <- ref_data_mistakes %>% distinct(spu_filename) %>% pull()


```

#### Apply References
Rerun the **Join Data & keys** sections above to update the `df_clean` dataframe.Apply your chosen references to actual spectral data to create the tidy dataframe `df_tidy` containing corrected sepectral reflectance values.

```{r apply_refs, echo=FALSE}
## Average 5 chosen ref measurements per DATE/SITE/BLOCK 
ref_summary <- ref_data %>% 
  ## The following steps expand the "Block" column to create one REF set per Block per Site. This structure is necessary for situtations where different refs are used for different blocks at the same site. 
  separate(Block, into = c("BX1", "BX2", "BX3", "BX4"), sep = ",") %>% #1: expand string entry in "Block" into separate columns -- NOTE: this step throws a "Warning: Expected 4 pieces." for sites w/less than 4 blocks
  gather(Block, BlockString, BX1:BX4) %>% #2: re-condense into one column, generates correct number of rows per site AND per block
  mutate(Block = str_squish(BlockString), BlockString=NULL) %>% #3: replace placeholder column names w/"B1-B4". Also removes whitespace from BlockString contents introduced by "separate" function
  filter(!is.na(Block)) %>% #4: remove empty rows for sites w/out B3 or B4
  ### Unnest Spectra & calculate
  unnest(Spectra) %>% 
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  mutate(CorrectionFactor = 1/Reflectance) %>% 

  ### The following code group repeated REF measurements, and takes the mean 
  group_by(Date,Site,Block,Wavelength) %>% 
  # group_by(Date,Site,Block,Wavelength, Integration_ms) %>% # to separate integration times
  summarize(ChA_REF = mean(ChA), ChB_REF = mean(ChB), CorrectionFactor = mean(ChA/ChB), int_REF = mean(Integration_ms), Notes_REF = str_c(Notes, collapse = "; "), ref_filenames = str_c(spu_filename,collapse = ", "))

## Join DATA with REFS
### Check no missing Spectra
spu_dataframe %>% filter(Spectra %>% map(is.null) %>% map_lgl(any))

spu_for_plots <- df %>% filter(!str_detect(Treatment, "REF|DARK")) %>% filter(!is.na(DateTime)) %>% 
  unnest(Spectra) %>% 
  filter(Wavelength > 400, Wavelength < 1000)


df_ref <- left_join(spu_for_plots, ref_summary) %>% 
  select(Date, DateTime, Site, Block, Treatment, Replicate, spu_filename, FileNum, Integration_ms, Weather, Notes, Notes_REF, ref_filenames, int_REF, Wavelength, ChB, ChA, ChB_REF, ChA_REF, CorrectionFactor) %>%
  mutate(raw_reflectance = ChB/ChA) %>% # the raw reflectance
  mutate(corrected_reflectance = raw_reflectance*CorrectionFactor) 


## Corrected Reflectances 
df_corrected <- df_ref %>% 
  nest(processed_spectra = c(Wavelength, ChB, ChA, ChB_REF, ChA_REF, CorrectionFactor, raw_reflectance, corrected_reflectance))
```

##### Check Application
```{r}
## Check all files have a corrected reflectance
corrected_spectra_files <- df_corrected %>% unnest(processed_spectra) %>% filter(!is.na(corrected_reflectance)) %>% select(spu_filename, Date, Site) %>% distinct()

## missing corrections
anti_join(df_corrected, corrected_spectra_files) %>% 
  group_by(Date, Site, Block, Treatment) %>% 
  summarize(Num_Files = n()) %>% print(n=100)

## Check that no important treatments are left out 
anti_join(df, df_corrected, by = "spu_filename") %>% pull(Treatment) %>% unique()
```

##### Plot Check
```{r}

(df_corrected %>% 
  #filter(Date == "2019-07-31", Site == "LMAT") %>%
  slice(1:10) %>% 
  unnest(processed_spectra) %>% 
  gather(key = Status, value = Reflectance, raw_reflectance, corrected_reflectance) %>% 
  
  # VIZ
  ggplot(mapping = aes(x = Wavelength, y = Reflectance )) + 
  geom_line(aes(color = spu_filename, linetype = Status)) + 
    facet_wrap(vars(Date, Site, Block, Treatment, FileNum)))

```


#### ? INTERPOLATION

### SAVE 
```{r}
processed_spu_data <- df_corrected %>% 
  select(spu_filename, DateTime, ref_filenames, processed_spectra) %>% unnest(processed_spectra) %>%
  select(spu_filename, DateTime, ref_filenames, Wavelength, ChB, ChA, raw_reflectance, CorrectionFactor, corrected_reflectance) %>% 
  nest(Spectra = c(Wavelength, ChB, ChA, raw_reflectance, CorrectionFactor, corrected_reflectance))

ref_spu_data <- ref_data %>% unnest(Spectra) %>% 
  select(spu_filename, DateTime, Wavelength, ChB, ChA, Reflectance) %>% 
  mutate(CorrectionFactor = 1/Reflectance, corrected_reflectance = NA) %>% 
  filter(Wavelength > 400, Wavelength < 1000) %>% 
  rename(raw_reflectance = Reflectance) %>% 
  nest(Spectra = c(Wavelength, ChB, ChA, raw_reflectance, CorrectionFactor, corrected_reflectance))

df_processed <- bind_rows(processed_spu_data, ref_spu_data)
  
## Save 
# processed_spu_filename <- paste0(data_path, dir_year, "_processed_spu_data.rds")
# 
# write_rds(df_processed, processed_spu_filename)
```




## Calculate INDICES

Currently, this only works for NDVI, EVI, and EVI2 as I haven't worked out spectral interpolation yet and the other indices need reflectance at a specific value (not a range). 

### Calculate Indices
```{r}

df_processed %>%  filter( Spectra %>% map(is.null) %>% map_lgl(any)) # remove files w/out spu_spectra

df_indices <- df %>% select(-Spectra) %>% 
  filter(!is.na(Treatment)) %>% 
  filter(!str_detect(Treatment, "REF|DARK")) %>% 
  inner_join(df_processed) %>% 
  ## Format for calculuating indices
  unnest(Spectra) %>% 
  select(-ChB, -ChA, -raw_reflectance, -CorrectionFactor) %>% 
  rename(Reflectance = corrected_reflectance) %>% 
  nest(Spectra = c(Wavelength, Reflectance)) %>%
  ## Calculate NDVI
  mutate(Indices = map(Spectra, function(x) calculate_indices(x, band_defns = band_defns, instrument = "MODIS", indices = c("NDVI", "EVI", "EVI2"))))
  
index_filename <- paste0(data_path, session_date, "_index_data.rds")
write_rds(df_indices, index_filename)


## convert to SHINY ready format

  
```


### Calculate INDICES for new spu files
```{r}
new_files <- spu_dataframe_new$spu_filename

df_processed_new <- df_processed %>% filter(spu_filename %in% new_files)

indices_new <- unispec_file_key_fix %>% 
  filter(!is.na(Treatment)) %>% 
  filter(!str_detect(Treatment, "REF|DARK")) %>% 
  inner_join(df_processed_new) %>% 
  ## Format for calculuating indices
  unnest(Spectra) %>% 
  select(-ChB, -ChA, -raw_reflectance, -CorrectionFactor) %>% 
  rename(Reflectance = corrected_reflectance) %>% 
  nest(Wavelength, Reflectance,.key = Spectra) %>%
  ## Calculate NDVI
  mutate(Indices = map(Spectra, function(x) calculate_indices(x, band_defns = band_defns, instrument = "MODIS", indices = c("NDVI", "EVI", "EVI2"))))

## SAVE UPDATED INDICES
indices_all <- bind_rows(df_indices, indices_new)
index_filename <- paste0(data_path, dir_year, "_index_data.rds")
write_rds(indices_all, index_filename)
  
```


### Visualize with other years 

### Check across years


## CHECKLIST/SUMMARY REPORT 

### Label Unispec Problems
```{r unispec_problem_key}

problem_files <- c(maxed_files, dim_files, absurd_files, missing_spu_files)

df_ref_filenames <- df_corrected %>% unnest(processed_spectra) %>% 
  select(spu_filename, ref_filenames) %>% distinct()

unispec_problem_key_pre <- unispec_file_key_fix%>% 
  mutate(Replicate = factor(Replicate)) %>% 
  # mislabeled 
  mutate(mislabeled = !is.na(key_fix)) %>% # mislabeled 
  # mismeasurement
  left_join(maxed_data) %>% # maxed_number, maxed_wavelenghs
  mutate(maxed = spu_filename %in% maxed_files) %>% 
  mutate(dim = spu_filename %in% dim_files) %>% # dim
  mutate(absurd_reflectance = spu_filename %in% c(zero_files, absurd_files)) %>% # absurd_reflectance
  # missing 
  mutate(missing_spu = spu_filename %in% missing_spu_files) %>%  # missing_spu
  left_join(df_ref_filenames)
  
# miscorrection 
ref_problem_key <- unispec_problem_key_pre %>%
  filter(spu_filename %in% ref_files) %>% # select ref files 
  #mutate(mislabeled = if_else(spu_filename %in% ref_mistakes, TRUE, mislabeled)) %>% 
  gather("problem", "status", mislabeled, maxed, dim, absurd_reflectance, missing_spu) %>% # row for each type of problem
  filter(!is.na(status)) %>% # remove non-problems
  filter(status != FALSE) %>% # remove non-problems
  group_by(spu_filename) %>% 
  select(spu_filename, problem, status) %>% 
  summarize(problems = str_c(unique(problem), collapse = ", ")) %>% # group: one row per spu_filename
  rename(ref_filenames = spu_filename, ref_problem = problems) # rename variables to join w/unispec_problem_key

## Unispec Problem Key
unispec_problem_key <- unispec_problem_key_pre %>% 
  separate_rows(ref_filenames, sep = ", ") %>% # split ref_filenames string into multiple rows per ref_filename
  left_join(ref_problem_key) %>% # adds ref_problem column 
  group_by(spu_filename) %>% 
  summarize(ref_problems = str_c(str_replace_na(ref_problem), collapse = "; ")) %>% # collapse to one row per spu_filename
  right_join(unispec_problem_key_pre) %>%  # add the rest of the key info back in
  mutate_at(.vars = c("ref_problems", "dim", "absurd_reflectance", "mislabeled", "missing_spu"), .funs = factor) %>% 
  mutate(file_problem = spu_filename %in% problem_files) %>% 
  select(spu_filename, Type, Date, Site, Block, Treatment, Replicate, Weather, Notes, missing_spu, file_problem, mislabeled, key_fix, maxed, maxed_number, maxed_wavelengths, dim, absurd_reflectance, ref_problems, ref_filenames) # order


### Summary table
unispec_problem_key %>%
  mutate_at(list(factor), .vars = vars(c("Site", "Block", "Treatment", "key_fix", "ref_problems"))) %>% 
  mutate(file_problem = spu_filename %in% problem_files) %>% 
  summary() 
```



### Upload Index Data
```{r}

index_data_2019 <- read_rds("/home/ruby/UnispecR/UnispecRecord/UnispecData/2019/2019_index_data.rds") %>% 
    unnest(Indices) %>% spread(Index, Value) %>% 
    mutate(DOY = lubridate::yday(DateTime))

## Make a character vector of all .spu raw files in folder
index_files <- list.files(path = "/home/ruby/UnispecData/2019_unprocessed_unispec_data/", pattern = "index_data.rds$", full.names = T, recursive=T)

# Read metadata text lines (9) from the spu files
index_filedata.shiny <- map_dfr(index_files, read_rds) %>%  select(DateTime, Site, Block, Treatment, Replicate, FileNum, Indices) %>% 
  unnest(Indices) %>% 
  
  # Spread Indices 
  group_by_at(vars(-Value)) %>%  # WEIRD WORK AROUND FOR SPREADING PROBLEM: group by everything other than the value column. 
  mutate(row_id=1:n()) %>% ungroup() %>%  # build group index
  spread(Index, Value) %>% 
  select(-row_id) %>%  # drop the index
  
  # Add the right dates 
  select(DateTime, Site, Block, Treatment, Replicate, FileNum, NDVI, EVI, EVI2) %>% 
  mutate(Year = lubridate::year(DateTime), Date = lubridate::date(DateTime), DOY = lubridate::yday(DateTime))

index_data_2019.shiny <- index_data_2019 %>% 
  select(DateTime, Site, Block, Treatment, Replicate, FileNum, NDVI, EVI, EVI2) %>% 
  mutate(Year = lubridate::year(DateTime), Date = lubridate::date(DateTime), DOY = lubridate::yday(DateTime))

index_data_2019.all.shiny <- bind_rows(index_data_2019.shiny, index_filedata.shiny) %>%  
  mutate(Block = as.numeric(str_extract(Block, "\\d"))) %>% 
  mutate(Replicate = as.character(Replicate))

index_data_update <- index_data %>% bind_rows(index_data_2019.all.shiny)

write_rds(index_data_update, "/home/ruby/UnispecR/Visualizations/shiny_unispec_cleaned/indices_2014-2019.updated.rds")

```
